### 2023-06-29 manuscript/manuscript.qmd
fix metrics and calculate n awakenings

---
title: Estimating Sleep Quality Metrics Using Free-Living Accelerometer Data From Thigh-Worn Devices in Comparison to an EEG-Based Sleep Tracking Device
author:
  - name: Esben Høegholm Lykke
    email: eskovgaard@health.sdu.dk
    affiliations: 
        - id: some-tech
          name: University of Southern Denmark
          department: Department of Sports Science and Clinical Biomechanics
          address: Campusvej 55
          city: Odense
          state: Denmark
          postal-code: 5230
    attributes:
        corresponding: true
  - name: Jan Christian Brønd
    email: jbrond@health.sdu.dk
    affiliations:
        - id: another-u
          name: University of Southern Denmark
          department: Department of Sports Science and Clinical Biomechanics
          address: Campusvej 55
          city: Odense
          state: Denmark
          postal-code: 5230
abstract: "LÆS IKKE! Det er volapyk på latin. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse vitae dictum eros, ullamcorper elementum orci. Sed laoreet nulla neque, pulvinar fermentum mi iaculis at. Nulla ultricies nibh sit amet vestibulum rutrum. Nam pharetra nisl sed ipsum maximus suscipit. Duis metus nunc, ullamcorper eu mi rutrum, tempus ultricies ante. Nunc vitae lectus nisi. Aliquam efficitur ut eros ut pellentesque. Aenean blandit, nisl nec efficitur interdum, nisi ipsum fermentum dolor, at tempus sem turpis in lacus. Curabitur sollicitudin lectus sit amet velit pellentesque laoreet. Ut posuere diam lobortis nisi eleifend tincidunt. Ut at euismod sem, sed dignissim ligula. Aliquam lacinia massa libero, id eleifend velit pulvinar ac. Fusce volutpat elit eu nulla viverra, nec tempus orci pellentesque."
keywords: 
  - Sleep
  - Accelerometry
  - EEG
  - Machine learning
  - Sleep quality metrics
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: npj Digital Medicine
      formatting: preprint
      model: 3p
      cite-style: super
    include-in-header: 
      text: '\usepackage{lineno}\linenumbers'
    fig-pos: 'b'
  docx:
    number-sections: true
editor: 
  markdown: 
    wrap: 72
---

# Introduction

An extensive array of research underlines the importance of sleep for
both mental and physical health[@ma2017; @meyer2022; @kpavlova2019;
@difrancesco2019]. Consequently, accurate sleep assessment methods are
essential for tracking sleep patterns, thereby enhancing our
comprehension of the sleep-health relationship. Furthermore, ensuring
high user-acceptability for these methods is essential in order to
conduct large-scale studies over prolonged periods.

While laboratory-based polysomnography is considered the gold standard
for objectively measuring sleep, its practicality in large-scale
epidemiological studies is limited due to its high cost, the need for
professional administration, and the substantial resources required for
specialized equipment[@vandewater2011]. As an alternative, diaries is
commonly used as low-cost and low-tech methods for sleep assessment in
population research. However, relying solely on diary-based methods may
introduce recall bias and other limitations[@moore2015]. A more feasible
approach in large-scale epidemiological studies is the use of
device-based measurement methods that can estimate sleep duration. This
method offers the advantage of being less burdensome for participants
and avoids potential biases associated with recall.

In this context, the Zmachine®️ Insight+ (ZM) emerges as a valuable
tool. Validated against polysomnography with favorable
results[@kaplan2014; @wang2015], the ZM provides comparable data without
the high costs or need for professional monitoring associated with
polysomnography. Furthermore, the ease of use of the ZM device makes it
compliant for free-living use[@pedersen2021]. This allows for the
analysis of multiple consecutive nights, as compared to single-night
recordings from PSG, thereby capturing the important variations in sleep
across multiple nights.

The introduction of body-worn accelerometers has provided an effective
and affordable alternative for objectively assessing sleep patterns in a
home environment over extended periods. These accelerometers collect
continuous, high-resolution data for several weeks without requiring
recharging, thus minimizing participant burden. Initial applications of
accelerometry for sleep and wake stage classification were based on
wrist movements. The original algorithm, developed in 1982 using simple
linear regression and validated with PSG [@webster1982], was later
refined in 1992 [@cole1992], leading to the widely used Cole-Kripke
model. Subsequent research on wrist-worn accelerometer data has employed
heuristic algorithms, advanced machine learning models, as well as
regression and deep learning techniques[@palotti2019; @cole1992;
@sazonov2004; @sadeh1994; @hees2015; @sundararajan2021].

Despite the well-developed field of accelerometer data analysis for
sleep detection from wrist and hip-worn devices, the same level of
advancement is not mirrored in studies utilizing thigh-worn
accelerometers. Methods for assessing sleep using wrist and hip-worn
accelerometers have greatly evolved over the years, employing an
extensive range of techniques. These include heuristic algorithms,
machine learning models, regression and deep learning techniques, all
tailored to the specific signal characteristics of wrist and hip-worn
devices [@palotti2019; @cole1992; @sazonov2004; @sadeh1994; @hees2015;
@sundararajan2021; @patterson_40_2023]. However, for thigh-worn
accelerometers, the landscape appears less mature, with only a handful
of studies investigating sleep detection algorithms. The majority of the
efforts are focused on delineating wakefulness from sleep, with
particular emphasis on the definition of 'waking time' and 'bedtime'
[@carlson2021; @inan-eroglu2021; @vanderberg2016; @winkler2016].
Furthermore, while strides have been made recently in estimating sleep
duration with these devices, with the introduction of a promising
algorithm and its comparison against PSG[@johansson_development_2023],
the field is still in its infancy when it comes to employing machine
learning techniques. Given the potential for accurate physical behavior
assessment that thigh-worn accelerometers
provide[@skotte_detection_2014; @arvidsson2019], a significant research
gap exists. Therefore, there is a pressing need for future studies to
develop techniques similar to those used for wrist and hip-worn
accelerometers, with the ultimate goal of establishing a more holistic,
accurate, and user-friendly method of sleep and physical activity
tracking.

Our primary objective in this study was to evaluate a range of machine
learning and deep learning models, utilizing the raw data collected from
a tri-axial thigh-worn accelerometer to estimate in-bed and sleep time.
To ensure the reliability and effectiveness of our models, we compared
their outputs with an EEG-based sleep tracking device, which we
considered the gold standard for measuring sleep. Furthermore, our
secondary goal was to assess the developed models' performance in
evaluating important sleep quality metrics, including sleep period time
(SPT), total sleep time (TST), sleep efficiency (SE), latency until
persistent sleep (LPS), and wake after sleep onset (WASO). ~~By
analyzing these additional metrics, we aimed to provide a comprehensive
evaluation of the model's capability in assessing various aspects of
sleep quality, allowing us to determine the overall effectiveness of our
models in accurately estimating different parameters related to sleep
duration and quality.~~

# Methods

## Dataset and participants

The current study leverages data from the SCREENS
project[@rasmussen2020], a study conducted from October 2018 to March
2019 in Middelfart, Southern Denmark, that evaluated the impact of
screen media usage on Danish families. For our analysis, we isolated
data from child participants between the ages of 6 to 10 years within
the SCREENS cohort. Our main sources of data were accelerometer readings
from Axivity AX3 devices attached to the children's thighs, and
electroencephalography (EEG) data derived from the ZM device. The
Axivity AX3, an unobtrusive 3-axis accelerometer, was positioned midway
between the hip and knee on the right anterior thigh, recording
participant movement data.

Sleep state information was extracted using the ZM, a product of General
Sleep Corporation. The ZM, utilizes advanced EEG hardware and signal
processing algorithms, employs three self-adhesive, disposable sensors
placed outside the hairline for reliable EEG signal acquisition. The ZM
uses two proprietary algorithms: Z-ALG and Z-PLUS. The Z-ALG is utilized
for accurate sleep detection, showcasing its aptness for in-home
monitoring[@kaplan2014], while the Z-PLUS effectively differentiates
sleep stages, as evidenced by its alignment with expert evaluations
using PSG data[@wang2015].

Finally, we affirm that the SCREENS study received approval from the
Regional Scientific Committee of Southern Denmark, and all data handling
processes complied with the General Data Protection Regulation (GDPR),
guaranteeing the ethical and secure management of participant
information.

## Data Preprocessing and Feature Extraction

In this study, raw data processing began with a low-pass filtration step
using a 4th order Butterworth filter with a 5 Hz cut-off frequency to
eliminate high-frequency noise. Following filtration, data were
partitioned into overlapping 2-second intervals, each successive
interval sharing a 50% overlap with the previous one similar to methods
described by Skotte et al.[@skotte_detection_2014]. Any non-wear data
was remove using previously described methods[@skovgaard2023] and data
was resampled to 30-second epochs so every sample classified by the
algorithms corresponds to a 30-second epoch scored during the ZM
recordings. Subsequently, we performed a feature extraction process that
yielded a set of 88 features, providing a robust characterization of the
data. Extracted from accelerometer and temperature signals, these
features include temporal elements that use both lag and lead values,
capturing dynamic data trends by incorporating measurements from
preceding and upcoming intervals. Furthermore, inspired by Walch et
al.[@walch2019], we developed sensor-independent features to encapsulate
circadian rhythms. These features offer unique insights not directly
discernible from sensor outputs and are meant to approximate the
changing drive of the circadian clock to sleep over the course of the
night (see @fig-sensor-independent). Furthermore, the feature set was
enriched by including signal characteristics, which encompass vector
magnitude, mean crossing rate, skewness, and kurtosis for each of the x,
y, and z dimensions. These features provide a comprehensive
understanding of the signal's attributes, facilitating a more nuanced
analysis of the underlying patterns. All features are summarized in
table ??? (overvej tabel til supp. material)

![Sensor-independent features of circadian rhythms across two
consecutive nights. A) cosinus feature, B) linear
feature.](visuals/sensor_independent.pdf){#fig-sensor-independent}

In addition to the engineered features, we chose to incorporate the
median-filtered raw predictions from the ZM device into our modeling
process. This decision stemmed from the understanding that children
typically undergo around five to eight sleep cycles per night, with
awakenings commonly occurring at the end of each
cycle[@galland_normal_2012]. Examining the raw ZM predictions, we noted
a significant overestimation in the number of awakenings per night for
the children in our study, exceeding what would be expected based on
typical sleep cycle patterns. Consequently, we elected to train and
evaluate our models using not only the raw ZM output, but also versions
that were subjected to 5-minute and 10-minute median filters. This
approach resulted in an anticipated, more age-appropriate count of
awakenings per night, providing a more accurate depiction of children's
sleep patterns (see @fig-zm-median).

![The difference in number of awakenings between the raw ZM predictions
vs. 5-minute, and 10-minute median filtered predictions for a random
night. Grey line is the raw predictions, black line is the median
filtered predictions. A: 5-minute median filter on raw ZM predictions,
B: 10-minute median filter on raw ZM
predictions.](visuals/zm_raw_vs_filtered.pdf){#fig-zm-median}

## Algorithms, Training and Validation

In our study, we utilize two distinct modeling strategies to predict
sleep patterns derived from thigh-mounted accelerometer data: a
sequential ensemble and a multiclass bidirectional Long Short-Term
Memory (biLSTM)[@hochreiter1997] neural network. The sequential ensemble
involves using four pairs of models, each pair consisting of the same
algorithm, that predict in-bed and sleep time successively. Conversely,
the biLSTM is a comprehensive model that predicts multiple classes
simultaneously, effectively handling complex temporal patterns. By
applying and comparing both, we aim to evaluate their relative
effectiveness.

### Models in Sequence

To predict in-bed time and sleep time accurately, we employed an
ensemble learning strategy based on sequential binary classification
models. This approach involved constructing a sequence of models using
multiple machine learning algorithms to improve predictive accuracy. The
process began with an initial model predicting in-bed time, followed by
a second model that utilized the output of the initial model to predict
sleep time. This sequential approach was applied across all four
algorithms detailed below, with each subsequent model leveraging the
outputs of the previous models for improved predictions.

1.  Logistic Regression (LREG): Logistic regression served as a simple
    and fast baseline model. However, due to its linear nature, it may
    struggle with capturing complex relationships and non-linear
    patterns present in the accelerometer data.

2.  Decision Tree (TREE): Decision trees are capable of handling
    non-linear patterns and are easily interpretable. However, they are
    prone to overfitting, particularly when dealing with complex
    patterns that require simultaneous consideration of multiple
    features.

3.  Single-layer Feed-forward Neural Network (SNN): Single-layer
    feed-forward neural networks can effectively capture non-linear
    relationships, even with their relatively simple structure. However,
    they tend to be more challenging to interpret compared to simpler
    models. Additionally, careful tuning of the network's architecture
    and training process is required to mitigate the risk of
    overfitting.

4.  XGBoost (XGB): XGBoost is a powerful algorithm known for its ability
    to provide highly accurate predictions and handle complex,
    non-linear patterns in the data. It also incorporates built-in
    methods to prevent overfitting. However, training XGBoost models can
    be computationally intensive, and interpreting the predictions it
    generates can pose challenges.

### Multiclass Model

The biLSTM is a multiclass classifier capable of predicting the classes
of out-bed-awake, in-bed-awake, and in-bed-asleep. The architecture of
the biLSTM consisted of four layers, each with 128 hidden units. The
choice of four layers and 128 hidden units in our model balanced
complexity and efficiency: it was deep enough to learn intricate
patterns yet feasible to train timely. Additionally, the bidirectional
nature of the LSTM, doubling the hidden units at each time step,
improved data comprehension and avoided overfitting.

The biLSTM model was chosen for sleep classification using accelerometer
data from a thigh-mounted device due to its advantages in modeling
temporal context, handling a large temporal scope, and enabling temporal
inference over any feature without the need for hand-designed temporal
features.

Previous studies have explored the use of LSTM models for sleep
detection, showing promising results in capturing complex temporal
patterns. Notable works by Sano et al. [@sano2019] and Chen et al.
[@chen2021] have demonstrated the potential of LSTM models in improving
sleep detection based on accelerometer data. However, further research
is necessary to fully comprehend the strengths and limitations of this
approach.

### Model Training

For the models in sequence, we trained four pairs of classification
models, each pair distinguishing between in-bed/out-of-bed and
asleep/awake states. The dataset was randomly split into a training set
(approximately 50% of the subjects) and a testing set (also
approximately 50%), ensuring that samples from the same subject were
never simultaneously used in both sets. Hyper-parameter optimization was
performed using 10-fold Monte Carlo cross-validation, with the F1 score
as the performance metric. The best-performing set of hyperparameters
was then used to fit the models to the full training dataset, maximizing
accuracy by leveraging all available data. Furthermore, the dataset for
predicting sleep was highly imbalanced which can pose challenges when
training machine learning models, as models may favor predicting the
majority class. To address this issue, we employed oversampling
techniques to ensure that each class had a roughly equal number of
training samples. Specifically, we utilized the Synthetic Minority
Over-sampling Technique (SMOTE)[@chawla2002], which generates new
samples by interpolating random samples with their nearest neighbors. In
our study, we implemented SMOTE using the themis R package[@themis],
resampling all classes to achieve a balanced distribution of training
samples.

The biLSTM model was trained using the Adam optimizer, which is
computationally efficient and adapts the learning rate during training.
Cross-entropy loss function was employed for its suitability in
multiclass classification with mutually exclusive classes. The softmax
activation function was selected for the output layer to obtain a
probability distribution over the classes. Data for training the biLSTM
were randomly divided into a training, validation, and test tensors
based on a 50/25/25 split. The model was evaluated using the F1 score on
both the training and validation sets. Early stopping was implemented
with a patience of 3 epochs, meaning training would stop if there was no
improvement in the validation loss for 3 consecutive epochs.

### Model Validation \[evt. bare "statistics"\]

In our study, we utilized standard evaluation metrics to assess the
performance of each model on an epoch-to-epoch basis. These include
accuracy ($accuracy = \frac{TP+TN}{TP+TN+FP+FN}$), sensitivity
($sensitivity = \frac{TP}{TP+FN}$), specificity
($specificity = \frac{TN}{TN+FP}$), precision
($precision = \frac{TP}{TP+FP}$), negative predictive value (NPV,
$NPV = \frac{TN}{TN + FN}$), and F1 score
($F_1 = 2 * \frac{precision * sensitivity}{precision + sensitivity}$).

In the context of our sequential learning strategy, the initial models
were tasked with the binary classification of in-bed vs. out-of-bed. For
this task, we assessed performance using the F1-score, accuracy,
sensitivity, specificity, and precision metrics. The second models in
our sequential learning strategy focused on the binary classification of
asleep vs. awake. For these models, we considered the same metrics, in
addition to the negative predictive rate. The class imbalance in this
case led us to compute the F1 score as an unweighted macro-average.
Additionally, we evaluated the multiclass classifier, biLSTM, using
macro-averaged F1-score, sensitivity, and precision. Furthermore, we
present precision-recall curves and receiver operating characteristic
curves, including the area under the curve, for each model and for each
type of ZM prediction.We considered both the in-bed/out-of-bed and
awake/asleep scoring tasks as binary classification problems,
designating in-bed and asleep as the positive labels and out-of-bed and
awake as the negative labels in accordance with previous
research[@hjorth2012; @kushida2001].

To evaluate our developed models' performance in assessing sleep quality
metrics, we employed Bland-Altman plots and Pearson correlations. These
methods were used to compare the sleep quality summaries calculated by
our models with those obtained from EEG-based ZM sleep quality
summaries. The sleep quality summaries included several metrics:

1.  Sleep Period Time (SPT) - This refers to the total duration of the
    sleep period, which is defined as the time from the start to the end
    of the ZM recording.

2.  Total Sleep Time (TST) - This is the time spent asleep within the
    SPT.

3.  Sleep Efficiency (SE) - This is the ratio between TST and SPT,
    representing the proportion of the sleep period that was actually
    spent asleep.

4.  Latency Until Persistent Sleep (LPS) - This metric represents the
    time it takes to transition from wakefulness to sustained sleep. It
    is calculated as the time from the beginning of the ZM recording
    until a period when 10 out of 12 minutes are scored as sleep.

5.  Wake After Sleep Onset (WASO) - This refers to the time spent awake
    after initially falling asleep and before the final awakening. In
    our analysis, a period is counted as 'awake' only if it consists of
    3 or more contiguous 30-second epochs.

R version 4.3.0 (2023-04-21)[@R-lang] and the Tidymodels[@tidymodels]
and Tidyverse[@tidyverse] suite of packages were used as the core tools
for model development and analyses. Python version
3.10.6[@10.5555/1593511] and PyTorch[@NEURIPS2019_9015] were used to
implement the biLSTM model. All code used to perform the analysis and
generate the figures in this paper are available in [this
repository](https://github.com/esbenlykke/sleep_study).

# Results

![Flowchart of eligible nights included in the study - revise text
boxes!](visuals/flowchart_of_elligible_nights.pdf)

The children children included in the analyses were 9.4 years old (SD =
2.1). The number of sleep epochs recorded in the study was 2,035,261,
which accounted for approximately 86% of the total sleep duration. This
metric indicates the granularity of sleep measurement, capturing the
discrete moments of sleep throughout the sleep period. Sleep epochs are
essential for analyzing sleep patterns and stages, as well as for
assessing sleep quality and efficiency.

As reported in @tbl-zm_overview the sleep quality metrics derived from
ZM predictions were modified by the implementation of 5-minute and
10-minute median filters. The Sleep Period Time (SPT) were consistent
across raw and filtered datasets (mean: 9.2 ± 2.1 hours) as this
corresponds to the length of the ZM recording. Total Sleep Time (TST)
and Sleep Efficiency (SE) increased in the filtered data, implying the
filters categorize some wakefulness as sleep. Specifically, TST
increased from a raw mean of 7.7 ± 1.9 hours to 8.1 ± 2.0 hours
(5-minute filter) and 8.2 ± 2.1 hours (10-minute filter), while SE rose
from 82.6 ± 12.0% to 86.4 ± 12.7% and 87.5 ± 12.9% respectively. Latency
to Persistent Sleep (LPS) also elevated, suggesting the filter smooths
out brief awakenings at sleep onset, leading to a prolonged time to
persistent sleep. The most significant change was seen in Wake After
Sleep Onset (WASO), which dropped from 39.0 ± 33.6 minutes in raw data
to 30.6 ± 46.8 minutes and 22.3 ± 55.4 minutes in the 5-minute and
10-minute filtered data, respectively. The results underscore the role
of median filters on sleep metrics.

```{r}
#| echo: false
#| message: false
#| label: tbl-zm_overview
#| tbl-cap: Overview of characteristics of the ZM sleep quality summaries. Values are represented as mean (SD).

source("code/table_overview_ZM_stats.R")

tbl_overview
```

## Performance on Epoch-To-Epoch Basis

The epoch-to-epoch evaluation of in-bed performance metrics, outlined in
@tbl-in_bed_performance, demonstrates practically equivalent performance
across all model types. While the Decision Tree model posted an F1 score
of 94.4% and accuracy of 95.3%, the Logistic Regression and Feed-Forward
Neural Network models each exhibited an F1 score of 95.0% and similar
accuracies of 95.7% and 95.8%, respectively. The XGBoost model, despite
recording the highest metrics with an F1 score of 95.4% and accuracy of
96.1%, outpaced the others only marginally. This underscores the
consistency of performance among these models in classifying in-bed
conditions.

```{r}
#| echo: false
#| message: false
#| label: tbl-in_bed_performance
#| tbl-cap: In-Bed Performance Metrics

source("code/table_in_bed_performance.R")

tbl_in_bed
```

@tbl-sleep_performance details the performance of the sequential models
on raw and median-filtered (5 and 10 minute) ZM predictions. The F1
Scores, which are unweighted macro averages, for raw ZM predictions
range from 71.05% to 76.18%. The models perform comparably, but the low
Specificity values (62.84% to 70.93%) suggest difficulty in correctly
classifying negative instances. Applying 5-minute median filtering
improves the performance metrics. The XGBoost model tops the charts with
an F1 Score of 79.22% and NPV of 74.00%. However, Specificity still
remains low, with values between 54.68% and 74.84% across all models.
With 10-minute median filtering, the metrics improve further. The
XGBoost model still leads with an F1 Score of 80.87% and an NPV of
75.76%. But, Specificity remains a concern, ranging from 57.47% to
76.35% across all models.

```{r}
#| echo: false
#| message: false
#| tbl-colwidths: auto
#| label: tbl-sleep_performance
#| tbl-cap: Sleep Performance Metrics

source("code/table_sleep_performance.R")

tbl_sleep_performance
```

The analysis of precision-recall and ROC curves across different models
and ZM prediction types shows varying performance. In terms of
precision-recall AUC, the Decision Tree model consistently outperforms
others, indicating its superior predictive accuracy (see
@fig-pr_curves). Conversely, the Neural Network model generally shows
weaker performance. However, for ROC AUC, the XGBoost model consistently
excels across all data types, indicating a strong ability to
differentiate between classes, while the Neural Network model tends to
underperform (see @fig-roc_curves). The F-measure (F1 score) shows
variable performance across different configurations but generally, the
Decision Tree model yields higher scores.

![Precision-recall curves of the models evaluated across the different
ZM predictions, including raw ZM predictions, as well as 5-minute and
10-minute median smoothing of the ZM raw predictions. The x-axis of the
plot represents the proportion of true wake epochs that were correctly
classified as wake, while the y-axis represents the proportion of all
epochs labeled as wake by the classifier that were classified correctly.
The area under the curve values are displayed as color-coded text in the
plot indicate the area under the Precision-Recall curve for each model
and condition.](visuals/plot_sleep_pr.pdf){#fig-pr_curves}

![Receiver operating characteristic curves of the models evaluated
across the different ZM predictions, including raw ZM predictions, as
well as 5-minute and 10-minute median smoothing of the ZM raw
predictions. The x-axis of the plot represents the proportion of true
asleep epochs that were incorrectly classified as awake, while the
y-axis represents the proportion of all epochs labeled as awake by the
classifier that were correctly classified. The area under the curve
values displayed are displayed as color-coded text in the plot to
indicate the area under the receiver operating characteristic curve for
each model and condition.](visuals/plot_sleep_roc.pdf){#fig-roc_curves}

The performance metrics of the biLSTM were...

\newpage

# References