### 2023-07-07 manuscript/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  super,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{lineno}\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{npj Digital Medicine}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Estimating Sleep Quality Metrics Using Free-Living Accelerometer Data From Thigh-Worn Devices in Comparison to an EEG-Based Sleep Tracking Device},
  pdfauthor={Esben Høegholm Lykke; Jan Christian Brønd},
  pdfkeywords={Sleep, Accelerometry, EEG, Machine learning, Sleep
quality},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Estimating Sleep Quality Metrics Using Free-Living Accelerometer
Data From Thigh-Worn Devices in Comparison to an EEG-Based Sleep
Tracking Device}
\author[1]{Esben Høegholm Lykke%
\corref{cor1}%
}
 \ead{eskovgaard@health.sdu.dk} 
\author[1]{Jan Christian Brønd%
%
}
 \ead{jbrond@health.sdu.dk}

\affiliation[1]{organization={University of Southern Denmark, Department
of Sports Science and Clinical Biomechanics},addressline={Campusvej
55},city={Odense},postcode={5230},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
LÆS IKKE! Det er volapyk på latin. Lorem ipsum dolor sit amet,
consectetur adipiscing elit. Suspendisse vitae dictum eros, ullamcorper
elementum orci. Sed laoreet nulla neque, pulvinar fermentum mi iaculis
at. Nulla ultricies nibh sit amet vestibulum rutrum. Nam pharetra nisl
sed ipsum maximus suscipit. Duis metus nunc, ullamcorper eu mi rutrum,
tempus ultricies ante. Nunc vitae lectus nisi. Aliquam efficitur ut eros
ut pellentesque. Aenean blandit, nisl nec efficitur interdum, nisi ipsum
fermentum dolor, at tempus sem turpis in lacus. Curabitur sollicitudin
lectus sit amet velit pellentesque laoreet. Ut posuere diam lobortis
nisi eleifend tincidunt. Ut at euismod sem, sed dignissim ligula.
Aliquam lacinia massa libero, id eleifend velit pulvinar ac. Fusce
volutpat elit eu nulla viverra, nec tempus orci pellentesque.
\end{abstract}

\begin{keyword}
    Sleep \sep Accelerometry \sep EEG \sep Machine learning \sep 
    Sleep quality
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

An extensive array of research underlines the importance of sleep for
both mental and physical
health\citep{ma2017, meyer2022, kpavlova2019, difrancesco2019}.
Consequently, accurate sleep assessment methods are essential for
tracking sleep patterns, thereby enhancing our comprehension of the
sleep-health relationship. Furthermore, ensuring high user-acceptability
for these methods is essential in order to conduct large-scale studies
over prolonged periods.

While laboratory-based polysomnography is considered the gold standard
for objectively measuring sleep, its practicality in large-scale
epidemiological studies is limited due to its high cost, the need for
professional administration, and the substantial resources required for
specialized equipment\citep{vandewater2011}. As an alternative, diaries
is commonly used as low-cost and low-tech methods for sleep assessment
in population research. However, relying solely on diary-based methods
may introduce recall bias and other limitations\citep{moore2015}. A more
feasible approach in large-scale epidemiological studies is the use of
device-based measurement methods that can estimate sleep duration. This
method offers the advantage of being less burdensome for participants
and avoids potential biases associated with recall.

In this context, the Zmachine®️ Insight+ (ZM) emerges as a valuable
tool. Validated against polysomnography with favorable
results\citep{kaplan2014, wang2015}, the ZM provides comparable data
without the high costs or need for professional monitoring associated
with polysomnography. Furthermore, the ease of use of the ZM device
makes it compliant for free-living use\citep{pedersen2021}. This allows
for the analysis of multiple consecutive nights, as compared to
single-night recordings from PSG, thereby capturing the important
variations in sleep across multiple nights.

The introduction of body-worn accelerometers has provided an effective
and affordable alternative for objectively assessing sleep patterns in a
home environment over extended periods. These accelerometers collect
continuous, high-resolution data for several weeks without requiring
recharging, thus minimizing participant burden. Initial applications of
accelerometry for sleep and wake stage classification were based on
wrist movements. The original algorithm, developed in 1982 using simple
linear regression and validated with PSG \citep{webster1982}, was later
refined in 1992 \citep{cole1992}, leading to the widely used Cole-Kripke
model. Subsequent research on wrist-worn accelerometer data has employed
heuristic algorithms, advanced machine learning models, as well as
regression and deep learning
techniques\citep{palotti2019, cole1992, sazonov2004, sadeh1994, hees2015, sundararajan2021}.

Despite the well-developed field of accelerometer data analysis for
sleep detection from wrist and hip-worn devices, the same level of
advancement is not mirrored in studies utilizing thigh-worn
accelerometers. Methods for assessing sleep using wrist and hip-worn
accelerometers have greatly evolved over the years, employing an
extensive range of techniques. These include heuristic algorithms,
machine learning models, regression and deep learning techniques, all
tailored to the specific signal characteristics of wrist and hip-worn
devices
\citep{palotti2019, cole1992, sazonov2004, sadeh1994, hees2015, sundararajan2021, patterson_40_2023}.
However, for thigh-worn accelerometers, the landscape appears less
mature, with only a handful of studies investigating sleep detection
algorithms. The majority of the efforts are focused on delineating
wakefulness from sleep, with particular emphasis on the definition of
`waking time' and `bedtime'
\citep{carlson2021, inan-eroglu2021, vanderberg2016, winkler2016}.
Furthermore, while strides have been made recently in estimating sleep
duration with these devices, with the introduction of a promising
algorithm and its comparison against
PSG\citep{johansson_development_2023}, the field is still in its infancy
when it comes to employing machine learning techniques. Given the
potential for accurate physical behavior assessment that thigh-worn
accelerometers provide\citep{skotte_detection_2014, arvidsson2019}, a
significant research gap exists. Therefore, there is a pressing need for
future studies to develop techniques similar to those used for wrist and
hip-worn accelerometers, with the ultimate goal of establishing a more
holistic, accurate, and user-friendly method of sleep and physical
activity tracking.

Our primary objective in this study was to evaluate a range of machine
learning and deep learning models, utilizing the raw data collected from
a tri-axial thigh-worn accelerometer to estimate in-bed and sleep time.
To ensure the reliability and effectiveness of our models, we compared
their outputs with an EEG-based sleep tracking device, which we
considered the gold standard for measuring sleep. Furthermore, our
secondary goal was to assess the developed models' performance in
evaluating important sleep quality parameters, including sleep period
time (SPT), total sleep time (TST), sleep efficiency (SE), latency until
persistent sleep (LPS), and wake after sleep onset (WASO).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{dataset-and-participants}{%
\subsection{Dataset and participants}\label{dataset-and-participants}}

The current study leverages data from the SCREENS
project\citep{rasmussen2020}, a study conducted from October 2018 to
March 2019 in Middelfart, Southern Denmark, that evaluated the impact of
screen media usage on Danish families. For our analysis, we focused on
data from child participants aged between 6 and 10 years within the
SCREENS cohort. Our primary sources of data were accelerometer readings
from Axivity AX3 devices attached to the children's thighs, and
electroencephalography (EEG) data derived from the ZM device. The
Axivity AX3, an unobtrusive 3-axis accelerometer, was positioned midway
between the hip and knee on the right anterior thigh, recording
participant movement data.

Sleep state information was extracted using the ZM, a product of General
Sleep Corporation. The ZM, which utilizes advanced EEG hardware and
signal processing algorithms, employs three self-adhesive, disposable
sensors placed outside the hairline for reliable EEG signal acquisition.
The ZM uses two proprietary algorithms: Z-ALG and Z-PLUS. The Z-ALG is
utilized for accurate sleep detection, showcasing its suitability for
in-home monitoring\citep{kaplan2014}, while the Z-PLUS effectively
differentiates sleep stages, as evidenced by its alignment with expert
evaluations using PSG data\citep{wang2015}. In the current study, we
treated all sleep stages as a single category effectively deducing the
output of the ZM to ``awake'' and ``asleep'' as the ability to
distinguish sleep stages are not a necessity to derive sleep quality
parameters of interest and to simplify the learning process of the
models.

Figure~\ref{fig-flow} illustrates the selection criteria applied to the
children's recordings from the SCREENS study. We included only ZM
recordings that were accompanied by complete accelerometer data and
lasted between 7 and 14 hours. Any night when the ZM reported sensor
issues was excluded. The children whose recordings were considered had
an average age of 9.4 years, with a standard deviation of 2.1. In their
raw form, the ZM predictions encompassed 696,779 epochs, each 30 seconds
long. Notably, approximately 84\% of the total ZM recording duration was
classified as sleep, resulting in an imbalance of the dataset.

\begin{figure}[b]

{\centering \includegraphics{visuals/flowchart_of_elligible_nights.pdf}

}

\caption{\label{fig-flow}Flowchart of eligible ZM recording nights
included in the study}

\end{figure}

Finally, we affirm that the SCREENS study received approval from the
Regional Scientific Committee of Southern Denmark, and all data handling
processes complied with the General Data Protection Regulation (GDPR),
ensuring the ethical and secure management of participant information.

\hypertarget{data-preprocessing-and-feature-extraction}{%
\subsection{Data Preprocessing and Feature
Extraction}\label{data-preprocessing-and-feature-extraction}}

In this study, data processing of the raw accelerometer data began with
a low-pass filtration step using a 4th order Butterworth filter with a 5
Hz cut-off frequency to eliminate high-frequency noise. Following
filtration, data were partitioned into overlapping 2-second intervals,
each successive interval sharing a 50\% overlap with the previous one
similar to methods described by Skotte et
al.\citep{skotte_detection_2014}. Any non-wear data was remove using
previously described methods\citep{skovgaard2023} and data was resampled
to 30-second epochs so every sample classified by the algorithms
corresponds to a 30-second epoch scored during the ZM recordings.
Subsequently, we performed a feature extraction process that yielded a
set of 88 features, providing a robust characterization of the data.
Extracted from accelerometer and temperature signals, these features
include temporal elements that use both lag and lead values, capturing
dynamic data trends by incorporating measurements from preceding and
upcoming intervals. Furthermore, inspired by Walch et
al.\citep{walch2019}, we incorporated sensor-independent features to
encapsulate circadian rhythms. These features offer unique insights not
directly discernible from sensor outputs and are meant to approximate
the changing drive of the circadian clock to sleep over the course of
the night (see Figure~\ref{fig-sensor-independent}). Furthermore, the
feature set was enriched by including signal characteristics, which
encompass vector magnitude, mean crossing rate, skewness, and kurtosis
for each of the x, y, and z dimensions. All features are summarized in
table ??? \textbf{in the supplementary matrials}. Subsequently, we
merged the ZM and corresponding accelerometer recordings. Any
overlapping time between the ZM and accelerometer data was treated as
`in-bed' time, with the remaining time considered `out-of-bed'. This
process yielded a dataset providing a around the clock temporal view of
each participant's activity and sleep patterns.

\begin{figure}[b]

{\centering \includegraphics{visuals/sensor_independent.pdf}

}

\caption{\label{fig-sensor-independent}Sensor-independent features of
circadian rhythms across two consecutive nights. A) cosinus feature, B)
linear feature.}

\end{figure}

In addition to the engineered features, we chose to incorporate the
median-filtered raw predictions from the ZM device into our modeling
process. This decision stemmed from the understanding that children
typically undergo around five to eight sleep cycles per night, with
awakenings most likely occurring at the end of each
cycle\citep{galland_normal_2012}. Upon examining the raw ZM predictions,
we noted a significant overestimation in the number of awakenings per
night for the children in our study, exceeding what would be expected
based on typical sleep cycle patterns. In particular, many of these
brief awakenings could be considered as noise, which when present in the
data, can potentially hinder the learning process of machine learning
models by obscuring the underlying patterns that the models are trying
to learn, leading to less accurate predictions. Consequently, we elected
to train and evaluate our models using not only the raw ZM output, but
also versions that were subjected to 5-minute and 10-minute median
filters. This approach, by mitigating this noise, resulted in an
anticipated, more age-appropriate count of awakenings per night,
providing a more accurate depiction of children's sleep patterns (see
Figure~\ref{fig-zm-median}).

\begin{figure}[b]

{\centering \includegraphics{visuals/zm_raw_vs_filtered.pdf}

}

\caption{\label{fig-zm-median}The difference in number of awakenings
between the raw ZM predictions vs.~5-minute, and 10-minute median
filtered predictions for a random night. Grey line is the raw
predictions, black line is the median filtered predictions. A: 5-minute
median filter on raw ZM predictions, B: 10-minute median filter on raw
ZM predictions.}

\end{figure}

\hypertarget{algorithms-training-and-validation}{%
\subsection{Algorithms, Training and
Validation}\label{algorithms-training-and-validation}}

In this study, we employed two distinct modeling strategies to analyze
sleep patterns from thigh-mounted accelerometer data. We used a
sequential strategy, comprising an ensemble of four pairs of models,
each pair featuring the same algorithm. This strategy aimed to make the
prediction task more manageable for the algorithms by breaking it down
into a sequence of two binary classifications: first predicting `in-bed'
time, then `sleep' time. Simultaneously, we also used a multiclass
approach utilizing a bidirectional Long Short-Term Memory
(biLSTM)\citep{hochreiter1997} neural network.

\hypertarget{models-in-sequence}{%
\subsubsection{Models in Sequence}\label{models-in-sequence}}

To predict in-bed time and sleep time accurately, we employed an
ensemble learning strategy based on sequential binary classification
models. This approach involved constructing a sequence of models using
multiple machine learning algorithms to improve predictive accuracy. The
process began with an initial model predicting in-bed time, followed by
a second model that utilized the output of the initial model to predict
sleep time. This sequential approach was applied across all four
algorithms detailed below, with each subsequent model leveraging the
outputs of the previous models for improved predictions.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Logistic Regression (LREG): Logistic regression served as a simple and
  fast baseline model. However, due to its linear nature, it may
  struggle with capturing complex relationships and non-linear patterns
  present in the accelerometer data.
\item
  Decision Tree (TREE): Decision trees are capable of handling
  non-linear patterns and are easily interpretable. However, they are
  prone to overfitting, particularly when dealing with complex patterns
  that require simultaneous consideration of multiple features.
\item
  Single-layer Feed-forward Neural Network (SNN): Single-layer
  feed-forward neural networks can effectively capture non-linear
  relationships, even with their relatively simple structure. However,
  they tend to be more challenging to interpret compared to simpler
  models. Additionally, careful tuning of the network's architecture and
  training process is required to mitigate the risk of overfitting.
\item
  XGBoost (XGB): XGBoost is a powerful algorithm known for its ability
  to provide highly accurate predictions and handle complex, non-linear
  patterns in the data. It also incorporates built-in methods to prevent
  overfitting. However, training XGBoost models can be computationally
  intensive, and interpreting the predictions it generates can pose
  challenges.
\end{enumerate}

\hypertarget{multiclass-model}{%
\subsubsection{Multiclass Model}\label{multiclass-model}}

In this study, we employed a biLSTM, a multiclass classifier, to predict
three distinct states: out-of-bed-awake, in-bed-awake, and
in-bed-asleep. The architecture of the biLSTM was set up with four
layers, each equipped with 128 hidden units. This configuration was
intentionally chosen to balance between model complexity and training
efficiency: it provided the depth necessary for learning intricate
patterns while remaining feasible for timely training. The bidirectional
design of the LSTM served to enhance data interpretation and mitigate
overfitting by doubling the hidden units at each time step. For input,
we used tensors shaped as sequences, with each sequence spanning 10
minutes and a step size of one.This approach follows in the footsteps of
previous studies that utilized LSTM models for sleep detection. These
studies showcased the promising potential of LSTMs in capturing complex
temporal patterns. Particularly, the works of Sano et al.
\citep{sano2019} and Chen et al. \citep{chen2021} demonstrated the
effectiveness of LSTM models in improving sleep detection using
accelerometer data, underscoring the value of this modeling approach.

\hypertarget{model-training}{%
\subsubsection{Model Training}\label{model-training}}

For the models in sequence, we trained four pairs of classification
models. Each pair was designed to distinguish between in-bed/out-of-bed
and asleep/awake states, respectively. The dataset was randomly split
into a training set and a testing set, each containing approximately
50\% of the subjects. This division ensured that samples from the same
night were never simultaneously present in both sets. To optimize
hyperparameters, we performed a 10-fold Monte Carlo cross-validation on
a regular grid, comprising 20 different combinations of hyperparameters.
The F1 score served as the optimization metric. The best-performing set
of hyperparameters was then used to fit the models to the full training
dataset. This approach allowed us to maximize performance by leveraging
all available data. Moreover, after extracting the in-bed time from the
initial sequential models, the imbalance on the resulting dataset could
cause biases during model training, as models may favor predicting the
majority class. To rectify this, we employed the Synthetic Minority
Over-sampling Technique (SMOTE)\citep{chawla2002}. SMOTE generates new
samples by interpolating random samples with their nearest neighbors. We
utilized the themis R package\citep{themis} to implement SMOTE,
resulting in a balanced distribution of training samples across both
classes.

The biLSTM model was trained to differentiate between three states:
out-of-bed-awake, in-bed-awake, and in-bed-asleep. The data used for
training the biLSTM was randomly divided into training, validation, and
test sets, based on a 50/25/25 split. We ensured that data from the same
night was not present across different sets. The model was trained using
the Adam optimizer, selected for its computational efficiency and
adaptability of the learning rate during training. Given the multiclass
classification task with mutually exclusive classes, we employed the
cross-entropy loss function. To obtain a probability distribution over
the classes, the softmax activation function was applied at the output
layer. We evaluated the model's performance using the F1 score on both
the training and validation sets. We implemented early stopping with a
patience of 3 epochs, halting the training process if there was no
improvement in the validation loss over three consecutive epochs.

\hypertarget{model-validation}{%
\subsubsection{Model Validation}\label{model-validation}}

In our study, we utilized standard evaluation metrics to assess the
performance of each model on an epoch-to-epoch basis. These include
accuracy (\(accuracy = \frac{TP+TN}{TP+TN+FP+FN}\)), sensitivity
(\(sensitivity = \frac{TP}{TP+FN}\)), specificity
(\(specificity = \frac{TN}{TN+FP}\)), precision
(\(precision = \frac{TP}{TP+FP}\)), negative predictive value (NPV,
\(NPV = \frac{TN}{TN + FN}\)), and F1 score
(\(F_1 = 2 * \frac{precision * sensitivity}{precision + sensitivity}\)).

In the context of our sequential learning strategy, the initial models
were tasked with the binary classification of in-bed vs.~out-of-bed. For
this task, we assessed performance using the F1-score, accuracy,
sensitivity, specificity, and precision metrics. The second models in
our sequential learning strategy focused on the binary classification of
asleep vs.~awake. For these models, we considered the same metrics, in
addition to the negative predictive rate. The class imbalance in this
case led us to compute the F1 score as an unweighted macro-average.
Additionally, we evaluated the multiclass classifier, biLSTM, using
macro-averaged F1-score, sensitivity, and precision. To further
illustrate model performance, we provide confusion matrices for the full
dataset, encompassing both in-bed and out-of-bed data. These matrices
report relative counts, column percentages (the proportion of the true
class accurately predicted), and row percentages (the proportion of
predictions correctly classified). We considered both the
in-bed/out-of-bed and awake/asleep scoring tasks as binary
classification problems, designating in-bed and asleep as the positive
labels and out-of-bed and awake as the negative labels in accordance
with previous research\citep{hjorth2012, kushida2001}.

To assess the performance of our models in deriving sleep quality
parameters, we utilized Bland-Altman plots and Pearson correlations. The
Bland-Altman method was employed specifically to determine the level of
agreement between two measurement techniques. Considering our dataset
contained multiple observations per subject, we integrated a bootstrap
procedure to address this extra source of variability. We calculated the
mean difference (bias) and defined the limits of agreement as the mean
difference plus or minus 1.96 times the standard deviation of these
differences. To ensure our measurements were robust and accounted for
intra-subject variability, we estimated the 95\% confidence intervals
for both the bias and the limits of agreement using a bias-corrected and
accelerated bootstrap method, utilizing 10,000 bootstrap replicates. The
sleep quality parameteres included the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Sleep Period Time (SPT) - This refers to the total duration of the
  sleep period, which is defined as the time from the start to the end
  of the ZM recording.
\item
  Total Sleep Time (TST) - This is the time spent asleep within the SPT.
\item
  Sleep Efficiency (SE) - This is the ratio between TST and SPT,
  representing the proportion of the sleep period that was actually
  spent asleep.
\item
  Latency Until Persistent Sleep (LPS) - This metric represents the time
  it takes to transition from wakefulness to sustained sleep. It is
  calculated as the time from the beginning of the ZM recording until a
  period when 10 out of 12 minutes are scored as sleep.
\item
  Wake After Sleep Onset (WASO) - This refers to the time spent awake
  after initially falling asleep and before the final awakening. In our
  analysis, a period is counted as `awake' only if it consists of 3 or
  more contiguous 30-second epochs which is also how the ZM summarizes
  WASO.
\end{enumerate}

R version 4.3.0 (2023-04-21)\citep{R-lang} and the
Tidymodels\citep{tidymodels} and Tidyverse\citep{tidyverse} suite of
packages were used as the core tools for model development and analyses.
Python version 3.10.6\citep{10.5555/1593511} and
PyTorch\citep{NEURIPS2019_9015} were used to implement the biLSTM model.
All code used to perform the analysis and generate the figures in this
paper are available in
\href{https://github.com/esbenlykke/sleep_study}{this repository}.

\hypertarget{results}{%
\section{Results}\label{results}}

As reported in Table~\ref{tbl-zm_overview} the sleep quality parameters
derived from ZM predictions were modified by the implementation of
5-minute and 10-minute median filters. SPT were consistent across raw
and filtered datasets (mean: 9.2 ± 2.1 hours), corresponding to the
length of the ZM recording. TST and SE increased in the filtered data,
implying the filters categorize some wakefulness as sleep. Specifically,
TST increased from a raw mean of 7.7 ± 1.9 hours to 8.1 ± 2.0 hours
(5-minute filter) and 8.2 ± 2.1 hours (10-minute filter), while SE rose
from 82.6 ± 12.0\% to 86.4 ± 12.7\% and 87.5 ± 12.9\% respectively. LPS
also elevated, suggesting the filter smooths out brief awakenings at
sleep onset, leading to a prolonged time to persistent sleep. A
significant change was seen in WASO, which dropped from 39.0 ± 33.6
minutes in raw data to 30.6 ± 46.8 minutes and 22.3 ± 55.4 minutes in
the 5-minute and 10-minute filtered data, respectively. The number of
awakenings was also considerably reduced with the application of
filters. In the raw data, the average number of awakenings was 34.46 ±
11.33 per night, which reduced to 4.43 ± 3.26 and 1.95 ± 2.01 for the
5-minute and 10-minute filtered data sets respectively.

\hypertarget{tbl-zm_overview}{}
\begin{longtable}{lrrrrrr}
\caption{\label{tbl-zm_overview}Overview of characteristics of the ZM sleep quality summaries per night.
Values are represented as mean (SD). }\tabularnewline

\toprule
 & SPT (hrs) & TST (hrs) & SE (\%) & LPS (min) & WASO (min) & Awakenings (N) \\ 
\midrule
Raw ZM Predictions & 9.2 (2.1) & 7.7 (1.9) & 82.6 (12) & 34.5 (27.9) & 39 (33.6) & 34.5 (11.3) \\ 
5-Min Median & 9.2 (2.1) & 8.1 (2) & 86.4 (12.7) & 36.3 (39.8) & 30.6 (46.8) & 4.4 (3.3) \\ 
10-Min Median & 9.2 (2.1) & 8.2 (2.1) & 87.5 (12.9) & 38 (48.7) & 22.3 (55.4) & 1.9 (2) \\ 
\bottomrule
\end{longtable}

\hypertarget{performance-on-epoch-to-epoch-basis}{%
\subsection{Performance on Epoch-to-Epoch
Basis}\label{performance-on-epoch-to-epoch-basis}}

The epoch-to-epoch evaluation of predicting in-bed time is outlined in
Table~\ref{tbl-in_bed_performance}, and demonstrates practically
equivalent performance across all model types. The F1 score ranges from
94.4\% (Decision Tree) to 95.4\% (XGBoost), while accuracy ranges from
95.3\% (Decision Tree) to 96.1\% (XGBoost). Sensitivity, Precision, and
Specificity also demonstrate consistent results across the models. The
XGBoost model, despite recording the highest metrics with an F1 score of
95.4\% and accuracy of 96.1\%, outpaced the others only marginally.

\hypertarget{tbl-in_bed_performance}{}
\begin{longtable}{lrrrrr}
\caption{\label{tbl-in_bed_performance}In-Bed Performance Metrics }\tabularnewline

\toprule
 & F1 Score (\%) & Accuracy (\%) & Sensitivity (\%) & Precision (\%) & Specificity (\%) \\ 
\midrule
Decision Tree & $94.4$ & $95.3$ & $93.1$ & $95.6$ & $96.9$ \\ 
Logistic Regression & $95.0$ & $95.7$ & $95.0$ & $94.9$ & $96.3$ \\ 
Feed-Forward Neural Net & $95.0$ & $95.8$ & $95.1$ & $95.0$ & $96.3$ \\ 
XGBoost & $95.4$ & $96.1$ & $95.8$ & $94.9$ & $96.2$ \\ 
\bottomrule
\end{longtable}

Table~\ref{tbl-sleep_performance} details the performance of the second
sequential models on raw and median-filtered (5 and 10 minute) ZM
predictions. These are the predictions of the sequential models' ability
to predict asleep/awake on only the extracted in-bed time. The F1
scores, which are unweighted macro averages, for raw ZM predictions
range from 71.05\% to 76.18\%. The models perform comparably, but the
low specificity values (62.84\% to 70.93\%) suggest difficulty in
correctly classifying awake epochs. Applying 5-minute median filtering
improves the performance metrics. The XGBoost model tops the charts with
an F1 score of 79.22\% and NPV of 74.00\%. However, specificity still
remains low, with values between 54.68\% and 74.84\% across all models.
With 10-minute median filtering, the metrics improve further. The
XGBoost model still leads with an F1 score of 80.87\% and an NPV of
75.76\%. But, specificity remains low, ranging from 57.47\% to 76.35\%
across all models.

\hypertarget{tbl-sleep_performance}{}
\begin{longtable}{lrrrrr}
\caption{\label{tbl-sleep_performance}Sleep Performance Metrics }\tabularnewline

\toprule
 & F1 Score (\%) & Precision (\%) & NPV (\%) & Sensitivity (\%) & Specificity (\%) \\ 
\midrule
\multicolumn{6}{l}{Raw ZM Predictions} \\ 
\midrule
Decision Tree & $72.9$ & $93.2$ & $48.4$ & $86.3$ & $67.1$ \\ 
Logistic Regression & $71.0$ & $93.7$ & $43.9$ & $82.7$ & $70.9$ \\ 
Neural Network & $71.8$ & $93.8$ & $45.1$ & $83.6$ & $70.8$ \\ 
XGBoost & $76.2$ & $92.8$ & $58.0$ & $91.3$ & $62.8$ \\ 
\midrule
\multicolumn{6}{l}{5-Min Median} \\ 
\midrule
Decision Tree & $75.5$ & $94.2$ & $55.5$ & $93.4$ & $59.0$ \\ 
Logistic Regression & $68.3$ & $95.8$ & $36.0$ & $81.4$ & $74.8$ \\ 
Neural Network & $71.7$ & $95.8$ & $41.6$ & $85.6$ & $73.1$ \\ 
XGBoost & $79.2$ & $93.9$ & $74.0$ & $97.3$ & $54.7$ \\ 
\midrule
\multicolumn{6}{l}{10-Min Median} \\ 
\midrule
Decision Tree & $76.3$ & $94.7$ & $58.1$ & $94.9$ & $57.5$ \\ 
Logistic Regression & $68.0$ & $96.5$ & $34.3$ & $81.9$ & $76.4$ \\ 
Neural Network & $71.0$ & $96.1$ & $39.5$ & $86.5$ & $71.4$ \\ 
XGBoost & $80.9$ & $94.9$ & $75.8$ & $97.7$ & $57.6$ \\ 
\bottomrule
\end{longtable}

The complete set of confusion matrices generated from data both
containing the out-of-bed and in-bed time are presented in
Figure~\ref{fig-conf_mat}. The figure shows favorable epoch-to-epoch
performance across across all sequential models, however, it is evident
that the biLSTM is less successful in classifying the in-bed-awake class
which cannot be deduced from the confusion matrices from the sequential
models. \textbf{{[}fejl i biLSTM confmat. Alle mats er ens{]}}

\begin{figure}[b]

{\centering \includegraphics{visuals/all_conf_mats.pdf}

}

\caption{\label{fig-conf_mat}Confusion matrices for binary sleep
prediction. The middle of each tile is the normalized count (overall
percentage) and, beneath it, the count. The bottom number is the column
percentage (target). At the right side of each tile is the row
percentage (prediction). i) decision tree, ii) logistic regression, iii)
feed-forward neural net, iv) xgboost, and v) biLSTM.}

\end{figure}

Table~\ref{tbl-biLSTM_performance} presents the performance of the
three-class biLSTM multiclassifier on raw and median-filtered (5-minute
and 10-minute) ZM predictions. Raw ZM predictions achieve F1 Scores
ranging from 71.36\% to 76.04\%, indicating overall good performance.
Applying 5-minute median filtering improves the metrics further,
resulting in F1 Scores ranging from 75.99\% to 78.53\%, demonstrating
enhanced precision and sensitivity. \textbf{{[}Overvej flere metrics.
Evt. per class\ldots{} pt er det lidt svært at sammenligne på tværs af
lstm og andre modeller ud over F1.{]}} With 10-minute median filtering,
F1 Scores range from 73.45\% to 73.93\%, maintaining good performance
but with limited information on Specificity. Further analysis is
required to assess performance across all classes.

\hypertarget{tbl-biLSTM_performance}{}
\begin{longtable}{cccccc}
\caption{\label{tbl-biLSTM_performance}Performance of the three-class biLSTM multicclassifier. }\tabularnewline

\toprule
 & F1 Score & Precision & NPV & Sensitivity & Specificity \\ 
\midrule
Raw ZM Predictions & $71.4$ & $76.0$ & $95.5$ & $70.4$ & $95.0$ \\ 
5-Min Median & $76.0$ & $78.5$ & $96.4$ & $74.6$ & $96.2$ \\ 
10-Min Median & $73.4$ & $73.1$ & $95.6$ & $73.9$ & $95.7$ \\ 
\bottomrule
\end{longtable}

\hypertarget{evaluation-of-sleep-quality-parameters}{%
\subsection{Evaluation of Sleep Quality
Parameters}\label{evaluation-of-sleep-quality-parameters}}

Table~\ref{tbl-ba_cor} presents a comparative analysis of the included
models used to predict various sleep quality parameters (SPT, TST, SE,
LPS, WASO) using the 5-minute median filtered ZM predictions. To see the
full table including models developed from raw ZM predictions and
10-minute median filtered ZM predictions, see \textbf{{[}SUPP. MAT.{]}}.
In terms of bias, the decision tree model consistently underestimated
SPT, TST, and SE, and overestimated LPS and WASO in comparison to ZM.
The logistic regression model had similar trends, with more pronounced
underestimation in TST and overestimation in LPS. The eed-forward neural
network also exhibited similar bias as the decision tree and the
logistic regression models, but with a higher overestimation in WASO. On
the other hand, the XGboost model showed least bias among all,
especially in its 5-minute median predictions. Considering LOA, the
decision tree had higher variability across different sleep quality
parameters and filtering techniques, particularly for LPS and WASO,
which indicates lower agreement with ZM. Other models had comparable LOA
but with notable exceptions. For example, TST LOA for the logistic
regression model was particularly wide in the 5-minute median
predictions. Correlation-wise, the pearson coefficient, revealed that
the XGboost model consistently had the highest correlation with ZM
across all sleep qualityparameters and filtering methods Notably, the
XGboost's 5-minute median predictions showed the strongest correlation
(0.66) for TST among all models and filtering techniques.

\hypertarget{tbl-ba_cor}{}
\begin{longtable}{lrrrr}
\caption{\label{tbl-ba_cor}Summary of Bias, Limits of Agreement (LOA), and Pearson Correlation for
various Sleep Parameter Predictions (SPT, TST, SE, LPS, WASO) using
different Machine Learning Models (Decision Tree, Logistic Regression,
Feed-Forward Neural Net, XGBoost) with Raw ZM Predictions, 5-Min and
10-Min Median as predictors. Each value is provided with its 95\%
Confidence Interval (CI). }\tabularnewline

\toprule
 & Bias (95\% CI) & LOA (95\% CI) & LOA (95\% CI) & Pearson, \emph{r} (95\% CI) \\ 
\midrule
\multicolumn{5}{l}{5-Min Median - Decision Tree} \\ 
\midrule
SPT (min) & -21.6 (-25.6;-17.6) & -117.5 (-125.6;-110.7) & 74.2 (63.9;85.9) & 0.54 (0.48;0.6) \\ 
TST (min) & -50.5 (-55.2;-46) & -161.4 (-175.8;-151.3) & 60.4 (51.5;71.7) & 0.48 (0.42;0.54) \\ 
SE (\%) & -5.5 (-6.3;-4.7) & -23.9 (-26.4;-22.2) & 12.9 (11.6;14.6) & 0.22 (0.14;0.29) \\ 
LPS (min) & 24.6 (19.7;29.1) & -88.8 (-115;-77.3) & 138 (126.2;156.7) & 0.06 (-0.02;0.14) \\ 
WASO (min) & 9.9 (6.5;14) & -79.4 (-109;-63.1) & 99.2 (80;136.1) & 0.15 (0.07;0.22) \\ 
\midrule
\multicolumn{5}{l}{5-Min Median - Logistic Regression} \\ 
\midrule
SPT (min) & -3.7 (-8;1) & -112.2 (-120.9;-105.2) & 104.8 (94;117.4) & 0.38 (0.3;0.44) \\ 
TST (min) & -139.7 (-146.9;-133) & -305.6 (-323.6;-291.8) & 26.2 (16.1;38.6) & 0.09 (0.01;0.17) \\ 
SE (\%) & -23.2 (-24.3;-22.2) & -48.1 (-50.9;-46.1) & 1.7 (0.1;3.8) & 0.13 (0.05;0.21) \\ 
LPS (min) & 58.1 (53.4;62.6) & -52.3 (-75;-40.1) & 168.6 (155.9;187.7) & 0.05 (-0.03;0.13) \\ 
WASO (min) & 45.4 (41.7;49.7) & -50.7 (-74.4;-38.4) & 141.5 (126.8;173) & 0.19 (0.11;0.27) \\ 
\midrule
\multicolumn{5}{l}{5-Min Median - Feed-Forward Neural Net} \\ 
\midrule
SPT (min) & -3.9 (-8.1;0.9) & -112.7 (-122;-105.2) & 104.9 (94.1;118.4) & 0.38 (0.3;0.44) \\ 
TST (min) & -126.5 (-132.8;-120.3) & -276.8 (-291.3;-264.7) & 23.9 (14.8;33.9) & 0.25 (0.17;0.32) \\ 
SE (\%) & -20.9 (-21.9;-19.9) & -44.3 (-46.3;-42.5) & 2.5 (1.1;4) & 0.21 (0.13;0.29) \\ 
LPS (min) & 35.3 (30.7;39.8) & -75.8 (-102.3;-63.4) & 146.5 (134.4;166.9) & 0.07 (-0.01;0.15) \\ 
WASO (min) & 45 (41.2;49.2) & -51.8 (-76.4;-39.1) & 141.7 (125.8;174.1) & 0.21 (0.14;0.29) \\ 
\midrule
\multicolumn{5}{l}{5-Min Median - XGboost} \\ 
\midrule
SPT (min) & 0.2 (-3.7;4.5) & -97.4 (-106.2;-90.3) & 97.8 (86.6;111) & 0.56 (0.5;0.61) \\ 
TST (min) & -7 (-10.8;-3.3) & -95.5 (-105.2;-88) & 81.4 (72.4;92.5) & 0.66 (0.61;0.7) \\ 
SE (\%) & -1.1 (-1.7;-0.5) & -15.6 (-17;-14.4) & 13.3 (12.2;14.7) & 0.44 (0.38;0.51) \\ 
LPS (min) & 28.5 (23.9;32.6) & -76.4 (-104.2;-63.3) & 133.4 (120.4;154.2) & 0.12 (0.04;0.2) \\ 
WASO (min) & -0.9 (-3.9;3) & -83.4 (-113.1;-66) & 81.7 (62;119.6) & 0.26 (0.18;0.33) \\ 
\midrule
\multicolumn{5}{l}{5-Min Median - biLSTM} \\ 
\midrule
SPT (min) & -36.1 (-41.7;-30) & -136.1 (-146.3;-126.9) & 64 (51.1;78.6) & 0.54 (0.45;0.62) \\ 
TST (min) & 12.8 (7.4;18.3) & -80.1 (-89.8;-72.3) & 105.8 (94.3;118.8) & 0.63 (0.55;0.69) \\ 
SE (\%) & 8 (7.2;8.8) & -5.1 (-6.8;-3.8) & 21.1 (19.5;23.1) & 0.16 (0.04;0.27) \\ 
LPS (min) & -15.7 (-25.9;-7.5) & -169 (-230.7;-127.9) & 137.6 (101.1;184.9) & 0.09 (-0.02;0.2) \\ 
WASO (min) & -3 (-9.9;7.7) & -144.1 (-197.2;-107.2) & 138.1 (90.8;211.4) & 0.02 (-0.1;0.13) \\ 
\bottomrule
\end{longtable}

Figure~\ref{fig-xgb_ba_cor} shows the agreement of the XGBoost model
trained on the 5-minute median filtered ZM predictions in estimating
sleep quality parameters as evaluated 5-minute median-smoothed ZM
derived sleep quality parameter. As can be seen on from the Bland-Altman
plots, a large portion of the nights are in almost perfect agreement
with the ZM and also a portion that can be considered outliers. This is
true for all sleep quality parameters.

\textbf{{[}Overvej at smide samtlige 75 plots i supp.mat.{]}}

\begin{figure}[b]

{\centering \includegraphics{visuals/median_5_xgboost_ba_cor.pdf}

}

\caption{\label{fig-xgb_ba_cor}Comparison of sleep quality parameters
derived from the XGboost model trained on the 5-minute smoothed ZM
predictions. The left column displays Bland-Altman plots. Dashed lines
represent the bias (the average difference between the two measurements)
and limits of agreement, with the 95\% confidence intervals represented
as the grayed areas. The right column displays scatter plots of
XGboost-derived vs ZM-derived sleep quality parameters. The dashed line
represents the identity line, while the full-drawn line represents the
best linear fit. Pearson's correlations are annotated in the upper left
corner.}

\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The present study embarked on an in-depth exploration of the utility and
comparative performance of different machine learning models in the
analysis of pediatric sleep data. The results offer promising insights
into the role of machine learning in enhancing the precision of sleep
quality metrics.

A significant finding was the impact of median filters on the sleep
metrics. Implementing 5-minute and 10-minute median filters led to
notable changes in key sleep parameters. In particular, the Total Sleep
Time (TST), Sleep Efficiency (SE), and Latency to Persistent Sleep (LPS)
increased, suggesting that the filters may categorize some instances of
wakefulness as sleep and smooth out brief awakenings. This underscores
the potential of median filters in refining sleep data and mitigating
the potential misclassification of sleep states.

Interestingly, the most marked change was observed in Wake After Sleep
Onset (WASO), with a substantial reduction in the filtered data. This
reduction, coupled with the decrease in the number of awakenings,
implies a significant role of median filters in sleep-wake cycle
characterization, potentially enhancing the reliability of sleep
assessment.

The machine learning models demonstrated high overall performance on an
epoch-to-epoch basis, with the XGBoost model marginally outperforming
the others. However, the models had more varied performance in
classifying sleep conditions, as evidenced by the sleep performance
metrics. The relatively lower Specificity values across all models
suggest a common challenge in correctly classifying negative instances.

In the precision-recall and ROC curve analyses, the Decision Tree model
showed superior precision-recall AUC, indicating strong predictive
accuracy. In contrast, the XGBoost model consistently demonstrated a
high ROC AUC, indicating a robust ability to differentiate between
classes. However, it's worth noting that the Neural Network model
generally showed weaker performance in these areas.

The three-class biLSTM multiclassifier showed overall good performance,
further supporting the potential of machine learning in sleep study.
However, the absence of Specificity values limits a complete assessment
of its performance, calling for further exploration in future studies.

It is unclear whether Johansson et
al.\citep{johansson_development_2023}{]} provide a method for detecting
the in-bed time\ldots{}

In conclusion, this study provides compelling evidence of the utility of
machine learning models and median filters in enhancing the precision
and reliability of pediatric sleep assessment. However, more research is
needed to overcome the challenge of correctly classifying negative
instances and to fully understand the performance of these models across
all classes and conditions. The findings underscore the potential of
machine learning in advancing sleep medicine, paving the way for more
accurate, reliable, and personalized sleep assessment in pediatric
populations.

\newpage

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}