### 2023-07-11 manuscript/manuscript.qmd
biLSTM performance as two binary.

---
title: "Improving Sleep Quality Estimation: A Comparative Study of Machine Learning and Deep Learning Techniques Utilizing Free-Living Accelerometer Data from Thigh-Worn Devices and EEG-Based Sleep Tracking"
author:
  - name: Esben Høegholm Lykke
    email: eskovgaard@health.sdu.dk
    affiliations: 
        - id: some-tech
          name: University of Southern Denmark
          department: Department of Sports Science and Clinical Biomechanics
          address: Campusvej 55
          city: Odense
          state: Denmark
          postal-code: 5230
    attributes:
        corresponding: true
  - name: Jan Christian Brønd
    email: jbrond@health.sdu.dk
    affiliations:
        - id: another-u
          name: University of Southern Denmark
          department: Department of Sports Science and Clinical Biomechanics
          address: Campusvej 55
          city: Odense
          state: Denmark
          postal-code: 5230
abstract: "LÆS IKKE! Det er volapyk på latin. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse vitae dictum eros, ullamcorper elementum orci. Sed laoreet nulla neque, pulvinar fermentum mi iaculis at. Nulla ultricies nibh sit amet vestibulum rutrum. Nam pharetra nisl sed ipsum maximus suscipit. Duis metus nunc, ullamcorper eu mi rutrum, tempus ultricies ante. Nunc vitae lectus nisi. Aliquam efficitur ut eros ut pellentesque. Aenean blandit, nisl nec efficitur interdum, nisi ipsum fermentum dolor, at tempus sem turpis in lacus. Curabitur sollicitudin lectus sit amet velit pellentesque laoreet. Ut posuere diam lobortis nisi eleifend tincidunt. Ut at euismod sem, sed dignissim ligula. Aliquam lacinia massa libero, id eleifend velit pulvinar ac. Fusce volutpat elit eu nulla viverra, nec tempus orci pellentesque."
keywords: 
  - Sleep
  - Accelerometry
  - EEG
  - Machine learning
  - Sleep quality
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: false
    journal:
      name: npj Digital Medicine
      formatting: preprint
      model: 3p
      cite-style: super
    include-in-header: 
      text: '\usepackage{lineno}\linenumbers'
    fig-pos: 'b'
  docx:
    number-sections: true
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: inline
---

# Introduction

A vast body of research highlights the critical role of sleep in
maintaining both mental and physical health[@ma2017; @meyer2022;
@kpavlova2019; @difrancesco2019]. Consequently, accurate sleep
assessment methods are crucial for tracking sleep patterns and improving
our understanding of the sleep-health relationship. Furthermore, the
ease of use and high acceptability of these methods are essential to
facilitate large-scale, longitudinal studies.

While laboratory-based polysomnography (PSG) is typically considered the
gold standard for objective sleep measurement, its practicality in
large-scale epidemiological studies is hindered due to high costs, the
necessity for professional administration, and it is also subject to
potential rater bias[@vandewater2011; @lee2022]. As an alternative,
diaries are commonly employed as cost-effective and low-tech methods for
sleep assessment in population research. However, reliance on
diary-based methods may lead to recall bias and other
limitations[@moore2015]. A more feasible approach in large-scale
epidemiological studies is to use device-based measurement methods that
can estimate sleep duration. This approach offers the advantage of being
less burdensome for participants and eliminates potential biases
associated with recall.

Body-worn accelerometers have emerged as an effective and affordable
alternative for objectively assessing sleep patterns at home over
extended periods. These devices gather continuous, high-resolution data
for several weeks without the need for recharging, thus reducing
participant burden. Initial applications of accelerometry for sleep and
wake stage classification were based on wrist movements, starting with
an algorithm developed in 1982 and validated with PSG [@webster1982].
This model was later refined in 1992 [@cole1992], giving rise to the
widely used Cole-Kripke model. As the field advanced, an array of
techniques, including heuristic algorithms, machine learning models,
regression, and deep learning, were employed to analyze wrist-worn
accelerometer data[@palotti2019; @cole1992; @sazonov2004; @sadeh1994;
@hees2015; @sundararajan2021].

While wrist and hip-worn devices have benefited from extensive
methodological development, thigh-worn accelerometers have not seen the
same level of advancement. Existing studies mainly focus on
distinguishing sleep from wakefulness, with emphasis on defining 'waking
time' and 'bedtime' [@carlson2021; @inan-eroglu2021; @vanderberg2016;
@winkler2016]. Recent strides in estimating sleep duration using these
devices have been made, including the introduction of a promising
algorithm and its comparison against PSG[@johansson_development_2023].
Despite these advancements, the application of machine learning
techniques in this area is still relatively unexplored. Considering the
potential of thigh-worn accelerometers for accurate physical behavior
assessment[@skotte_detection_2014; @arvidsson2019], there is a
significant research gap. Therefore, future studies need to develop
techniques similar to those used for wrist and hip-worn accelerometers,
with the ultimate goal of establishing a more holistic, accurate, and
user-friendly method of sleep and physical activity tracking.

The Zmachine®️ Insight+ (ZM) emerges as a valuable tool within this
landscape. Favorably validated against PSG[@kaplan2014; @wang2015], the
ZM provides comparable data without the high costs or the need for
professional monitoring typically associated with PSG. Crucially, the ZM
facilitates multi-night analysis in free-living conditions due to its
ease of use[@pedersen2021], capturing the natural variations in sleep
patterns. This makes it advantageous over single-night PSG, particularly
as a gold standard data source in machine learning tasks, as it provides
multiple nights of measurements without inter-rater bias. Despite these
benefits, the ZM, like PSG, still poses a significant participant burden
and cost, reinforcing the need for more accessible alternatives like
accelerometers.

Our primary objective in this study was to evaluate a range of machine
learning and deep learning models, utilizing the raw data collected from
a tri-axial thigh-worn accelerometer to estimate in-bed and sleep time.
To ensure the reliability and effectiveness of our models, we compared
their outputs with an EEG-based sleep tracking device, which we
considered the gold standard for measuring sleep. Furthermore, our
secondary goal was to assess the developed models' performance in
evaluating important sleep quality parameters, including sleep period
time (SPT), total sleep time (TST), sleep efficiency (SE), latency until
persistent sleep (LPS), and wake after sleep onset (WASO).

# Methods

## Dataset and participants

The current study leverages data from the SCREENS
project[@rasmussen2020], a study conducted from October 2018 to March
2019 in Middelfart, Southern Denmark, that evaluated the impact of
screen media usage on Danish families. For our analysis, we focused on
data from child participants aged between 6 and 10 years within the
SCREENS cohort. Our primary sources of data were accelerometer readings
from Axivity AX3 devices attached to the children's thighs, and
electroencephalography (EEG) data derived from the ZM device. The
Axivity AX3, an unobtrusive 3-axis accelerometer, was positioned midway
between the hip and knee on the right anterior thigh, recording
participant movement data.

Sleep state information was extracted using the ZM, a product of General
Sleep Corporation. The ZM, which utilizes advanced EEG hardware and
signal processing algorithms, employs three self-adhesive, disposable
sensors placed outside the hairline for reliable EEG signal acquisition.
The ZM uses two proprietary algorithms: Z-ALG and Z-PLUS. The Z-ALG is
utilized for accurate sleep detection, showcasing its suitability for
in-home monitoring[@kaplan2014], while the Z-PLUS effectively
differentiates sleep stages, as evidenced by its alignment with expert
evaluations using PSG data[@wang2015]. In the current study, we treated
all sleep stages as a single category effectively deducing the output of
the ZM to "awake" and "asleep" as the ability to distinguish sleep
stages are not a necessity to derive sleep quality parameters of
interest and to simplify the learning process of the models.

@fig-flow illustrates the selection criteria applied to the children's
recordings from the SCREENS study. We included only ZM recordings that
were accompanied by complete accelerometer data and lasted between 7 and
14 hours. Any night when the ZM reported sensor issues was excluded. The
children whose recordings were considered had an average age of 9.4
years, with a standard deviation of 2.1. In their raw form, the ZM
predictions encompassed 696,779 epochs, each 30 seconds long. Notably,
approximately 84% of the total ZM recording duration was classified as
sleep, resulting in an imbalance of the dataset.

![Flowchart of eligible ZM recording nights included in the
study](visuals/flowchart_of_elligible_nights.pdf){#fig-flow}

Finally, we affirm that the SCREENS study received approval from the
Regional Scientific Committee of Southern Denmark, and all data handling
processes complied with the General Data Protection Regulation (GDPR),
ensuring the ethical and secure management of participant information.

## Data Preprocessing and Feature Extraction

In this study, data processing of the raw accelerometer data began with
a low-pass filtration step using a 4th order Butterworth filter with a 5
Hz cut-off frequency to eliminate high-frequency noise. Following
filtration, data were partitioned into overlapping 2-second intervals,
each successive interval sharing a 50% overlap with the previous one
similar to methods described by Skotte et al.[@skotte_detection_2014].
Any non-wear data was remove using previously described
methods[@skovgaard2023] and data was resampled to 30-second epochs so
every sample classified by the algorithms corresponds to a 30-second
epoch scored during the ZM recordings. Subsequently, we performed a
feature extraction process that yielded a set of 88 features, providing
a robust characterization of the data. Extracted from accelerometer and
temperature signals, these features include temporal elements that use
both lag and lead values, capturing dynamic data trends by incorporating
measurements from preceding and upcoming intervals. Furthermore,
inspired by Walch et al.[@walch2019], we incorporated sensor-independent
features to encapsulate circadian rhythms. These features offer unique
insights not directly discernible from sensor outputs and are meant to
approximate the changing drive of the circadian clock to sleep over the
course of the night (see @fig-sensor-independent). Furthermore, the
feature set was enriched by including signal characteristics, which
encompass vector magnitude, mean crossing rate, skewness, and kurtosis
for each of the x, y, and z dimensions. All features are summarized in
table ??? **in the supplementary matrials**. Subsequently, we merged the
ZM and corresponding accelerometer recordings. Any overlapping time
between the ZM and accelerometer data was treated as 'in-bed' time, with
the remaining time considered 'out-of-bed'. This process yielded a
dataset providing a around the clock temporal view of each participant's
activity and sleep patterns.

![Sensor-independent features of circadian rhythms across two
consecutive nights. A) cosinus feature, B) linear
feature.](visuals/sensor_independent.pdf){#fig-sensor-independent}

In addition to the engineered features, we chose to incorporate the
median-filtered raw predictions from the ZM device into our modeling
process. This decision stemmed from the understanding that children
typically undergo around five to eight sleep cycles per night, with
awakenings most likely occurring at the end of each
cycle[@galland_normal_2012]. Upon examining the raw ZM predictions, we
noted a significant overestimation in the number of awakenings per night
for the children in our study, exceeding what would be expected based on
typical sleep cycle patterns. In particular, many of these brief
awakenings could be considered as noise, which when present in the data,
can potentially hinder the learning process of machine learning models
by obscuring the underlying patterns that the models are trying to
learn, leading to less accurate predictions. Consequently, we elected to
train and evaluate our models using not only the raw ZM output, but also
versions that were subjected to 5-minute and 10-minute median filters.
This approach, by mitigating this noise, resulted in an anticipated,
more age-appropriate count of awakenings per night, providing a more
accurate depiction of children's sleep patterns (see @fig-zm-median).

![The difference in number of awakenings between the raw ZM predictions
vs. 5-minute, and 10-minute median filtered predictions for a random
night. Grey line is the raw predictions, black line is the median
filtered predictions. A: 5-minute median filter on raw ZM predictions,
B: 10-minute median filter on raw ZM
predictions.](visuals/zm_raw_vs_filtered.pdf){#fig-zm-median}

## Algorithms, Training and Validation

In this study, we employed two distinct modeling strategies to analyze
sleep patterns from thigh-mounted accelerometer data. We used a
sequential strategy, comprising an ensemble of four pairs of models,
each pair featuring the same algorithm. This strategy aimed to make the
prediction task more manageable for the algorithms by breaking it down
into a sequence of two binary classifications: first predicting 'in-bed'
time, then 'sleep' time. Simultaneously, we also used a multiclass
approach utilizing a bidirectional Long Short-Term Memory
(biLSTM)[@hochreiter1997] neural network.

### Models in Sequence

To predict in-bed time and sleep time accurately, we employed an
ensemble learning strategy based on sequential binary classification
models. This approach involved constructing a sequence of models using
multiple machine learning algorithms to improve predictive accuracy. The
process began with an initial model predicting in-bed time, followed by
a second model that utilized the output of the initial model to predict
sleep time. This sequential approach was applied across all four
algorithms detailed below, with each subsequent model leveraging the
outputs of the previous models for improved predictions.

1.  Logistic Regression (LREG): Logistic regression served as a simple
    and fast baseline model. However, due to its linear nature, it may
    struggle with capturing complex relationships and non-linear
    patterns present in the accelerometer data.

2.  Decision Tree (TREE): Decision trees are capable of handling
    non-linear patterns and are easily interpretable. However, they are
    prone to overfitting, particularly when dealing with complex
    patterns that require simultaneous consideration of multiple
    features.

3.  Single-layer Feed-forward Neural Network (SNN): Single-layer
    feed-forward neural networks can effectively capture non-linear
    relationships, even with their relatively simple structure. However,
    they tend to be more challenging to interpret compared to simpler
    models. Additionally, careful tuning of the network's architecture
    and training process is required to mitigate the risk of
    overfitting.

4.  XGBoost (XGB): XGBoost is a powerful algorithm known for its ability
    to provide highly accurate predictions and handle complex,
    non-linear patterns in the data. It also incorporates built-in
    methods to prevent overfitting. However, training XGBoost models can
    be computationally intensive, and interpreting the predictions it
    generates can pose challenges.

### Multiclass Model

We also employed a biLSTM, a multiclass classifier, to predict three
distinct states: out-of-bed-awake, in-bed-awake, and in-bed-asleep. The
architecture of the biLSTM was set up with four layers, each equipped
with 128 hidden units. This configuration was intentionally chosen to
balance between model complexity and training efficiency: it provided
the depth necessary for learning intricate patterns while remaining
feasible for timely training. The bidirectional design of the LSTM
served to enhance data interpretation and mitigate overfitting by
doubling the hidden units at each time step. For input, we used tensors
shaped as sequences, with each sequence spanning 10 minutes and a step
size of one.This approach follows in the footsteps of previous studies
that utilized LSTM models for sleep detection. These studies showcased
the promising potential of LSTMs in capturing complex temporal patterns.
Particularly, the works of Sano et al. [@sano2019] and Chen et al.
[@chen2021] demonstrated the effectiveness of LSTM models in improving
sleep detection using accelerometer data, underscoring the value of this
modeling approach.

### Model Training

For the models in sequence, we trained four pairs of classification
models. Each pair was designed to distinguish between in-bed/out-of-bed
and asleep/awake states, respectively. The dataset was randomly split
into a training set and a testing set, each containing approximately 50%
of the subjects. This division ensured that samples from the same night
were never simultaneously present in both sets. To optimize
hyperparameters, we performed a 10-fold Monte Carlo cross-validation on
a regular grid, comprising 20 different combinations of hyperparameters.
The F1 score served as the optimization metric. The best-performing set
of hyperparameters was then used to fit the models to the full training
dataset. This approach allowed us to maximize performance by leveraging
all available data. Moreover, after extracting the in-bed time from the
initial sequential models, the imbalance on the resulting dataset could
cause biases during model training, as models may favor predicting the
majority class. To rectify this, we employed the Synthetic Minority
Over-sampling Technique (SMOTE)[@chawla2002]. SMOTE generates new
samples by interpolating random samples with their nearest neighbors. We
utilized the themis R package[@themis] to implement SMOTE, resulting in
a balanced distribution of training samples across both classes.

The biLSTM model was trained to differentiate between three states:
out-of-bed-awake, in-bed-awake, and in-bed-asleep. The data used for
training the biLSTM was randomly divided into training, validation, and
test sets, based on a 50/25/25 split. We ensured that data from the same
night was not present across different sets. The model was trained using
the Adam optimizer, selected for its computational efficiency and
adaptability of the learning rate during training. Given the multiclass
classification task with mutually exclusive classes, we employed the
cross-entropy loss function. To obtain a probability distribution over
the classes, the softmax activation function was applied at the output
layer. We evaluated the model's performance using the F1 score on both
the training and validation sets. We implemented early stopping with a
patience of 3 epochs, halting the training process if there was no
improvement in the validation loss over three consecutive epochs.

### Model Validation

In our study, we utilized standard evaluation metrics to assess the
performance of each model on an epoch-to-epoch basis. These include
accuracy ($accuracy = \frac{TP+TN}{TP+TN+FP+FN}$), sensitivity
($sensitivity = \frac{TP}{TP+FN}$), specificity
($specificity = \frac{TN}{TN+FP}$), precision
($precision = \frac{TP}{TP+FP}$), negative predictive value (NPV,
$NPV = \frac{TN}{TN + FN}$), and F1 score
($F_1 = 2 * \frac{precision * sensitivity}{precision + sensitivity}$).

In the context of our sequential learning strategy, the initial models
were tasked with the binary classification of in-bed vs. out-of-bed. For
this task, we assessed performance using the F1-score, accuracy,
sensitivity, specificity, and precision metrics. The second models in
our sequential learning strategy focused on the binary classification of
asleep vs. awake. For these models, we considered the same metrics, in
addition to the negative predictive rate. The class imbalance in this
case led us to compute the F1 score as an unweighted macro-average.
Additionally, we evaluated the multiclass classifier, biLSTM, using the
same metrics. To do this, we considered the multiclass output as to
binary classifications, where the first was out-of-bed vs the rest and
the second binary classification as in-bed-awake vs in-bed-asleep. To
further illustrate model performance, we provide confusion matrices for
the full dataset, encompassing both in-bed and out-of-bed data. These
matrices report relative counts, column percentages (the proportion of
the true class accurately predicted), and row percentages (the
proportion of predictions correctly classified). We considered both the
in-bed/out-of-bed and awake/asleep scoring tasks as binary
classification problems, designating in-bed and asleep as the positive
labels and out-of-bed and awake as the negative labels in accordance
with previous research[@hjorth2012; @kushida2001].

To assess the performance of our models in deriving sleep quality
parameters, we utilized Bland-Altman plots and Pearson correlations. The
Bland-Altman method was employed specifically to determine the level of
agreement between two measurement techniques. Considering our dataset
contained multiple observations per subject, we integrated a bootstrap
procedure to address this extra source of variability. We calculated the
mean difference (bias) and defined the LOA as the mean difference plus
or minus 1.96 times the standard deviation of these differences. To
ensure our measurements were robust and accounted for intra-subject
variability, we estimated the 95% confidence intervals for both the bias
and the LOA using a bias-corrected and accelerated bootstrap method,
utilizing 10,000 bootstrap replicates. The sleep quality parameteres
included are defined as follews in accordance with the ZM definitions:

1.  Sleep Period Time (SPT) - This refers to the total duration of the
    sleep period, which is defined as the time from the start to the end
    of the ZM recording.

2.  Total Sleep Time (TST) - This is the time spent asleep within the
    SPT.

3.  Sleep Efficiency (SE) - This is the ratio between TST and SPT,
    representing the proportion of the sleep period that was actually
    spent asleep.

4.  Latency Until Persistent Sleep (LPS) - This metric represents the
    time it takes to transition from wakefulness to sustained sleep. It
    is calculated as the time from the beginning of the ZM recording
    until the first period when 10 out of 12 minutes are scored as
    sleep.

5.  Wake After Sleep Onset (WASO) - This refers to the time spent awake
    after initially falling asleep and before the final awakening. In
    our analysis, a period is counted as 'awake' only if it consists of
    3 or more contiguous 30-second epochs which is also how the ZM
    summarizes WASO.

R version 4.3.0 (2023-04-21)[@R-lang] and the Tidymodels[@tidymodels]
and Tidyverse[@tidyverse] suite of packages were used as the core tools
for model development and analyses. Python version
3.10.6[@10.5555/1593511] and PyTorch[@NEURIPS2019_9015] were used to
implement the biLSTM model. All code used to perform the analysis and
generate the figures in this paper are available in [this
repository](https://github.com/esbenlykke/sleep_study).

# Results

As reported in @tbl-zm_overview the sleep quality parameters derived
from ZM predictions were modified by the implementation of 5-minute and
10-minute median filters. SPT were consistent across raw and filtered
datasets (mean: 9.2 ± 2.1 hours), corresponding to the length of the ZM
recording. TST and SE increased in the filtered data, implying the
filters categorize some wakefulness as sleep. Specifically, TST
increased from a raw mean of 7.7 ± 1.9 hours to 8.1 ± 2.0 hours
(5-minute filter) and 8.2 ± 2.1 hours (10-minute filter), while SE rose
from 82.6 ± 12.0% to 86.4 ± 12.7% and 87.5 ± 12.9% respectively. LPS
also elevated, suggesting the filter smooths out brief awakenings at
sleep onset, leading to a prolonged time to persistent sleep. A
significant change was seen in WASO, which dropped from 39.0 ± 33.6
minutes in raw data to 30.6 ± 46.8 minutes and 22.3 ± 55.4 minutes in
the 5-minute and 10-minute filtered data, respectively. The number of
awakenings was also considerably reduced with the application of
filters. In the raw data, the average number of awakenings was 34.46 ±
11.33 per night, which reduced to 4.43 ± 3.26 and 1.95 ± 2.01 for the
5-minute and 10-minute filtered data sets respectively.

```{r}
#| echo: false
#| message: false
#| label: tbl-zm_overview
#| tbl-cap: Overview of characteristics of the ZM sleep quality summaries per night. Values are represented as mean (SD).

source("code/table_overview_ZM_stats.R")

tbl_overview
```

## Performance on Epoch-to-Epoch Basis

The epoch-to-epoch evaluation of predicting in-bed time is outlined in
@tbl-in_bed_performance, and demonstrates practically equivalent
performance across all model types. The F1 score ranges from 94.4%
(Decision Tree) to 95.4% (XGBoost), while accuracy ranges from 95.3%
(Decision Tree) to 96.1% (XGBoost). Sensitivity, Precision, and
Specificity also demonstrate consistent results across the models. The
XGBoost model, despite recording the highest metrics with an F1 score of
95.4% and accuracy of 96.1%, outpaced the others only marginally.

```{r}
#| echo: false
#| message: false
#| label: tbl-in_bed_performance
#| tbl-cap: In-bed performance metrics

source("code/table_in_bed_performance.R")

tbl_in_bed
```

@tbl-sleep_performance details the performance of all sequential model
types on raw and median-filtered (5 and 10 minute) ZM predictions for
sleep/wake classification. For raw ZM predictions, the F1 scores, which
are unweighted macro averages, range from 65.6% (biLSTM) to 76.2%
(XGBoost). All models perform comparably, but the low specificity values
(62.5% to 70.9%) suggest difficulty in correctly classifying awake
epochs. Applying a 5-minute median filter improves the performance
metrics. The XGBoost model tops the charts with an F1 score of 79.2% and
NPV of 74.0%. However, specificity still remains low, with values
between 54.7% (XGBoost) and 74.8% (Logistic Regression) across all
models. With a 10-minute median filter, the metrics improve further. The
XGBoost model still leads with an F1 score of 80.9% and an NPV of 75.9%.
But, specificity remains low, ranging from 57.5% (Decision Tree) to
76.4% (Logistic Regression) across all models.

```{r}
#| echo: false
#| message: false
#| tbl-colwidths: auto
#| label: tbl-sleep_performance
#| tbl-cap: Performance of the sleep/wake classification of the sequential models.

source("code/table_sleep_performance.R")

tbl_sleep_performance
```

A complete set of confusion matrices generated from data both containing
the out-of-bed and in-bed time are presented in @fig-conf_mat. The
figure shows favorable epoch-to-epoch performance across across all
sequential models, however, it is evident that the biLSTM is less
successful in classifying the in-bed-awake class which cannot be deduced
from the confusion matrices from the sequential models.

![Confusion matrices for binary sleep prediction. The middle of each
tile is the normalized count (overall percentage) and, beneath it, the
count. The bottom number is the column percentage (target). At the right
side of each tile is the row percentage (prediction). i) decision tree,
ii) logistic regression, iii) feed-forward neural net, iv) XGBoost, and
v) biLSTM.](visuals/all_conf_mats.pdf){#fig-conf_mat}

## Evaluation of Sleep Quality Parameters

@tbl-ba_cor presents a comparative analysis of the included models used
to predict various sleep quality parameters (SPT, TST, SE, LPS, WASO)
using the 5-minute median filtered ZM predictions. To see the full table
including models developed from raw ZM predictions and 10-minute median
filtered ZM predictions, see **\[SUPP. MAT.\]**. In terms of bias, the
decision tree model consistently underestimated SPT, TST, and SE, and
overestimated LPS and WASO in comparison to ZM. The logistic regression
model had similar trends, with more pronounced underestimation in TST
and overestimation in LPS. The eed-forward neural network also exhibited
similar bias as the decision tree and the logistic regression models,
but with a higher overestimation in WASO. On the other hand, the XGBoost
model showed least bias among all, especially in its 5-minute median
predictions. Considering LOA, the decision tree had higher variability
across different sleep quality parameters and filtering techniques,
particularly for LPS and WASO, which indicates lower agreement with ZM.
Other models had comparable LOA but with notable exceptions. For
example, TST LOA for the logistic regression model was particularly wide
in the 5-minute median predictions. Correlation-wise, the pearson
coefficient, revealed that the XGBoost model consistently had the
highest correlation with ZM across all sleep qualityparameters and
filtering methods Notably, the XGBoost's 5-minute median predictions
showed the strongest correlation (0.66) for TST among all models and
filtering techniques.

```{r}
#| echo: false
#| message: false
#| label: tbl-ba_cor
#| tbl-cap: Summary of Bias, Limits of Agreement, and Pearson Correlation for various Sleep Parameter Predictions (SPT, TST, SE, LPS, WASO) using different Machine Learning Models (Decision Tree, Logistic Regression, Feed-Forward Neural Net, XGBoost) with Raw ZM Predictions, 5-Min and 10-Min Median as predictors. Each value is provided with its 95% Confidence Interval (CI).

source("code/table_ba_cor_statistics.R")

short_tbl_ba_cor

```

@fig-xgb_ba_cor shows the agreement between the XGBoost model, trained
on 5-minute median filtered ZM predictions, and the 5-minute
median-smoothed ZM-derived sleep quality parameters. The Bland-Altman
plot for the SPT reveals a significant level of agreement with the ZM,
as evidenced by a positive correlation. However, the presence of extreme
outliers widens the limits of agreement (LOA). The scatterplot for SPT
also demonstrates a positive trend, indicating a moderate linear
correlation between the XGBoost model and the ZM-derived sleep quality
parameters.In terms of TST, perfect agreement is not observed for a
substantial number of nights, but there is a positive correlation
between the XGBoost model and ZM-derived sleep quality parameters. The
bias and LOA for TST are comparable to those observed for SPT,
indicating a consistent level of agreement between the two methods. The
scatterplot for TST also shows a slightly higher correlation, primarily
driven by the absence of extreme outliers.Furthermore, the remaining
three sleep quality parameters, SE, LPS, and WASO, exhibit
heteroscedasticity in contrast to SPT and TST. This outcome is expected
as achieving 100% sleep efficiency is relatively rare, resulting in less
disagreement between the methods as values approach the upper limit.
However, as sleep efficiency decreases, the potential for discrepancies
and differing interpretations between the methods increases, leading to
greater heteroscedasticity. A moderate linear correlation is observed
between the XGBoost model and ZM-derived sleep quality parameters for
SE, indicating a positive relationship. However, a poor correlation is
observed for LPS and WASO, suggesting less agreement between the two
methods for these parameters. Similar plots for all models are available
in the supplementary materials.

![Comparison of sleep quality parameters derived from the XGBoost model
trained on the 5-minute smoothed ZM predictions. The left column
displays Bland-Altman plots. Dashed lines represent the bias (the
average difference between the two measurements) and LOA, with the 95%
confidence intervals represented as the grayed areas. The right column
displays scatter plots of XGBoost-derived vs ZM-derived sleep quality
parameters. The dashed line represents the identity line, while the
full-drawn line represents the best linear fit. Pearson's correlations
are annotated in the upper left
corner.](visuals/median_5_xgboost_ba_cor.pdf){#fig-xgb_ba_cor}

# Discussion

In an effort to mature the methods for estimating sleep from thigh-worn
accelerometers, we evaluated various models for predicting in-bed and
sleep time and their derived sleep quality parameters. Furthermore, we
trained and evaluated the models using raw and median-filtered
gold-standard predictions from the ZM. In general, all sequential models
performed well at predicting in-bed time. More challenging was it to
distinguish wake from sleep on the extracted in-bed time, however, the
performance of the sequential models were enhanced by the application of
median filters. Moreover, even though the multiclass biLSTM showed good
performance across F1 score, precision and NPV, the derived sleep
quality parameters were not on par with the XGBoost model which
demonstrated the highest performance metrics across all evaluations,
including epoch-to-epoch prediction and sleep quality parameter
derivatives. Despite this, all sequential models showed low specificity
values, indicating difficulty in correctly classifying awake epochs
during time in bed. The application of 5-minute and 10-minute median
filters improved the performance metrics of all models, with the XGBoost
model consistently leading. The filters increased total sleep time and
sleep efficiency, while reducing wake after sleep onset and the number
of awakenings. The XGBoost model also showed the least bias and highest
correlation with ZM across sleep quality parameters.

Limited research exists regarding the epoch-to-epoch effectiveness of
classifying in-bed time based on data from thigh-worn accelerometers.
Nevertheless, Carlson and colleagues provided compelling insights. They
demonstrated that a third-party algorithm, "ProcessingPal," and a
proprietary one, "CREA," achieved accuracies of 91% and 86%
respectively. These algorithms, evaluated against self-reported
measures[@carlson2021], produced F1 scores as high as 95% and 96%. These
figures are consistent with the performance of our sequential models,
which also achieved F1 scores and accuracy scores exceeding 95% in
identifying in-bed time. In our study, in-bed time is equated with SPT.
All models, with the exception of XGBoost, underestimated SPT. The
biLSTM model showed the greatest underestimation, with a bias of -36
minutes, reflecting trends observed in previous research. Winkler et al.
developed an algorithm that, despite a strong correlation (Pearson
correlation coefficient = .67) between their algorithmic results and
diary-recorded waking times, overestimated waking wear time by more than
30 minutes, resulting in an underestimation of in-bed
time[@winkler2016]. This trend was further confirmed when Inan-Eroglu et
al. examined Winkler et al.'s algorithm, revealing a underestimation of
9.8 minutes in bed time compared to self-reported
measures[@inan-eroglu2021]. In contrast, a study by van der Berg et al.
reported a slight underestimation of in-bed time. They employed a unique
approach with their algorithm, which relied on quantifying the number
and duration of sedentary periods to determine time in bed, and active
periods (standing or stepping) to identify wake times[@vanderberg2016].
Finally. it is important to note that high predictive performance in
determining in-bed time does not necessarily translate to accurate
predictions of broader sleep quality parameters. The crucial task of
detecting awake periods during in-bed time, a key factor in assessing
sleep quality, may not be effectively captured by in-bed time
predictions alone. Indeed, underestimating in-bed time could result in
overestimating waking time during in-bed time. Furthermore, the
distinction between actual sleep and time spent in bed, often overlooked
but vital in sleep research, is critical for a comprehensive
understanding of sleep quality.

To the best of our knowledge, Johansson and
colleagues[@johansson_development_2023] are the only researchers who
have reported epoch-to-epoch performance metrics for sleep scoring using
thigh-worn accelerometers, beyond just "waking time" and "in-bed time."
They achieved a mean sensitivity of 0.84, specificity of 0.55, and
accuracy of 0.80, using a single-night evaluation dataset of 71
subjects. Despite our models achieving a sensitivity above 97%, they,
like Johansson et al.'s algorithm, struggled with detecting in-bed awake
epochs. This is reflected in the low specificity scores, ranging from
54.7% to 76.4%, reported in our study. The challenge of low specificity
is not unique to methods using thigh-worn devices. Conley et al.'s
meta-analysis[@conley2019] reported similar findings when estimating
sleep using wrist-worn accelerometers among healthy adults, with a mean
sensitivity, accuracy, and specificity of 0.89, 0.88, and 0.53,
respectively. These figures align with our results regarding in-bed
time. Patterson and colleagues[@patterson_40_2023] recently summarized
the performance of various heuristic algorithms, machine learning, and
deep learning models used to predict sleep. They found the mean
sensitivity and specificity to be 93% (SD = 2.8) and 60% (SD = 11.1)
respectively. These findings underscore the challenge of automating the
detection of in-bed awake periods. Interestingly, despite low
specificity values for most of our models and configurations, we
observed an overestimation of LPS and WASO, contrasting with most
previous research[@conley2019; @palotti2019]. This overestimation of
wake epochs is evident from the low NPV scores, indicating that only a
small proportion of the wake predictions are actually correct. This
discrepancy may be driven by the SMOTE process used to balance the
dataset. If the synthetic "wake" examples created by SMOTE are not
representative of the true "wake" data, the models might learn to
incorrectly classify certain "sleep" epochs as "wake". This could lead
to an overestimation of LPS and WASO, as the models are incorrectly
identifying more periods of wakefulness during the sleep period.

often times imbalanced datasets only containing night recordings

The sleep quality parameter that our models consistently corresponded
most with the ZM derived sleep quality parameters were SPT and TST

something on outliers...maybe related to incl 180 degrees problem.

single night vs multiple nights/days

It is unclear whether Johansson et al.[@johansson_development_2023]\]
provide a method for detecting the in-bed time which hinders its
applicability for researchers that want to ...

A significant finding was the impact of median filters on the sleep
metrics. Implementing 5-minute and 10-minute median filters led to
notable changes in key sleep parameters. In particular, the Total Sleep
Time (TST), Sleep Efficiency (SE), and Latency to Persistent Sleep (LPS)
increased, suggesting that the filters may categorize some instances of
wakefulness as sleep and smooth out brief awakenings. This underscores
the potential of median filters in refining sleep data and mitigating
the potential misclassification of sleep states.

Interestingly, the most marked change was observed in Wake After Sleep
Onset (WASO), with a substantial reduction in the filtered data. This
reduction, coupled with the decrease in the number of awakenings,
implies a significant role of median filters in sleep-wake cycle
characterization, potentially enhancing the reliability of sleep
assessment.

The machine learning models demonstrated high overall performance on an
epoch-to-epoch basis, with the XGBoost model marginally outperforming
the others. However, the models had more varied performance in
classifying sleep conditions, as evidenced by the sleep performance
metrics. The relatively lower Specificity values across all models
suggest a common challenge in correctly classifying negative instances.

In the precision-recall and ROC curve analyses, the Decision Tree model
showed superior precision-recall AUC, indicating strong predictive
accuracy. In contrast, the XGBoost model consistently demonstrated a
high ROC AUC, indicating a robust ability to differentiate between
classes. However, it's worth noting that the Neural Network model
generally showed weaker performance in these areas.

The three-class biLSTM multiclassifier showed overall good performance,
further supporting the potential of machine learning in sleep study.
However, the absence of Specificity values limits a complete assessment
of its performance, calling for further exploration in future studies.

strengths and limitations

ZM as gold-standard? Without a gold standard, both methods can both
contribute to disagreement.

external validation dataset/generalization

expand the feature set (Palotti)

In conclusion, this study provides compelling evidence of the utility of
machine learning models and median filters in enhancing the precision
and reliability of pediatric sleep assessment. However, more research is
needed to overcome the challenge of correctly classifying negative
instances and to fully understand the performance of these models across
all classes and conditions. The findings underscore the potential of
machine learning in advancing sleep medicine, paving the way for more
accurate, reliable, and personalized sleep assessment in pediatric
populations.

\newpage

# References