### 2023-07-14 manuscript/manuscript.qmd
first draft of manuscript done.

---
title: "Improving Sleep Quality Estimation: A Comparative Study of Machine Learning and Deep Learning Techniques Utilizing Free-Living Accelerometer Data from Thigh-Worn Devices and EEG-Based Sleep Tracking"
author:
  - name: Esben Høegholm Lykke
    email: eskovgaard@health.sdu.dk
    affiliations: 
        - id: some-tech
          name: University of Southern Denmark
          department: Department of Sports Science and Clinical Biomechanics
          address: Campusvej 55
          city: Odense
          state: Denmark
          postal-code: 5230
    attributes:
        corresponding: true
  - name: Jan Christian Brønd
    email: jbrond@health.sdu.dk
    affiliations:
        - id: another-u
          name: University of Southern Denmark
          department: Department of Sports Science and Clinical Biomechanics
          address: Campusvej 55
          city: Odense
          state: Denmark
          postal-code: 5230
abstract: "LÆS IKKE! Det er volapyk på latin. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse vitae dictum eros, ullamcorper elementum orci. Sed laoreet nulla neque, pulvinar fermentum mi iaculis at. Nulla ultricies nibh sit amet vestibulum rutrum. Nam pharetra nisl sed ipsum maximus suscipit. Duis metus nunc, ullamcorper eu mi rutrum, tempus ultricies ante. Nunc vitae lectus nisi. Aliquam efficitur ut eros ut pellentesque. Aenean blandit, nisl nec efficitur interdum, nisi ipsum fermentum dolor, at tempus sem turpis in lacus. Curabitur sollicitudin lectus sit amet velit pellentesque laoreet. Ut posuere diam lobortis nisi eleifend tincidunt. Ut at euismod sem, sed dignissim ligula. Aliquam lacinia massa libero, id eleifend velit pulvinar ac. Fusce volutpat elit eu nulla viverra, nec tempus orci pellentesque."
keywords: 
  - Sleep
  - Accelerometry
  - EEG
  - Machine learning
  - Sleep quality
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: false
    journal:
      name: npj Digital Medicine
      formatting: preprint
      model: 3p
      cite-style: super
    include-in-header: 
      text: '\usepackage{lineno}\linenumbers'
    fig-pos: 'b'
  docx:
    number-sections: true
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: inline
---

# Introduction

A vast body of research highlights the critical role of sleep in
maintaining both mental and physical health[@ma2017; @meyer2022;
@kpavlova2019; @difrancesco2019]. Consequently, accurate sleep
assessment methods are crucial for tracking sleep patterns and improving
our understanding of the sleep-health relationship. Furthermore, the
ease of use and high acceptability of these methods are essential to
facilitate large-scale, longitudinal studies.

The traditional gold standard for objective sleep measurement,
laboratory-based polysomnography (PSG), has been found to be impractical
in large-scale epidemiological studies due to its high cost, need for
professional administration, and susceptibility to rater
bias[@vandewater2011; @lee2022]. As an alternative, diaries have been
used due to their cost-effectiveness and simplicity, although they are
subject to recall bias and other limitations[@moore2015]. An innovative
approach involves device-based measurement methods. These tools, which
estimate sleep duration, are advantageous due to their reduced
participant burden and elimination of potential recall biases. A
prominent example of such tools is body-worn accelerometers, which offer
a practical and affordable means of objectively assessing sleep patterns
at home for extended periods. Accelerometers collect continuous,
high-resolution data for several weeks without requiring recharging,
further minimizing participant burden. Their use in sleep and wake
classification began with a wrist movement-based algorithm developed in
1982, and validated using PSG[@webster1982]. This algorithm was refined
in 1992[@cole1992], leading to the widely adopted Cole-Kripke model.
With advancements in the field, a variety of techniques, including
heuristic algorithms, machine learning models, regression, and deep
learning, are now used to analyze data from hip and wrist-worn
accelerometers[@palotti2019; @cole1992; @sazonov2004; @sadeh1994;
@hees2015; @sundararajan2021].

While wrist and hip-worn devices have benefited from extensive
methodological development, thigh-worn accelerometers have not seen the
same level of advancement. Existing studies mainly focus on
distinguishing sleep from wakefulness, with emphasis on defining 'waking
time' and 'bedtime' [@carlson2021; @inan-eroglu2021; @vanderberg2016;
@winkler2016]. Recent strides in estimating sleep duration using these
devices have been made, including the introduction of a promising
algorithm and its comparison against PSG[@johansson_development_2023].
Despite these advancements, the application of machine learning
techniques in this area is still unexplored. Considering the potential
of thigh-worn accelerometers for accurate physical behavior
assessment[@skotte_detection_2014; @arvidsson2019], there is a
significant research gap. Therefore, future studies need to develop
techniques similar to those used for wrist and hip-worn accelerometers,
with the ultimate goal of establishing a more holistic, accurate, and
user-friendly method of sleep and physical activity tracking.

The Zmachine®️ Insight+ (ZM) emerges as a valuable tool within this
landscape. Favorably validated against PSG[@kaplan2014; @wang2015], the
ZM provides comparable data without the high costs or the need for
professional monitoring typically associated with PSG. Crucially, the ZM
facilitates multi-night analysis in free-living conditions due to its
ease of use[@pedersen2021], capturing the natural variations in sleep
patterns. This makes it advantageous over single-night PSG, particularly
as a gold standard data source in machine learning tasks, as it provides
multiple nights of measurements without inter-rater bias. Despite these
benefits, the ZM, like PSG, still poses a significant participant burden
and cost, reinforcing the need for more accessible alternatives like
accelerometers.

Our primary objective in this study was to evaluate a range of machine
learning and deep learning models, utilizing the raw data collected from
a tri-axial thigh-worn accelerometer to estimate in-bed and sleep time.
To ensure the reliability and effectiveness of our models, we compared
their outputs with an EEG-based sleep tracking device, which we, in this
current study, considered as the gold standard for measuring sleep.
Furthermore, our secondary goal was to assess the developed models'
performance in evaluating important sleep quality parameters, including
sleep period time (SPT), total sleep time (TST), sleep efficiency (SE),
latency until persistent sleep (LPS), and wake after sleep onset (WASO).

# Methods

## Dataset and participants

The current study leverages data from the SCREENS
project[@rasmussen2020], a study conducted from October 2018 to March
2019 in Middelfart, Southern Denmark, that evaluated the impact of
screen media usage on Danish families. For our analysis, we focused on
data from child participants aged between 6 and 10 years within the
SCREENS cohort. Our primary sources of data were accelerometer readings
from Axivity AX3 devices attached to the children's thighs, and
electroencephalography data derived from the ZM device. The Axivity AX3,
an unobtrusive 3-axis accelerometer, was positioned midway between the
hip and knee on the right anterior thigh, recording participant movement
data.

Sleep state information was extracted using the ZM, a product of General
Sleep Corporation. The ZM, which utilizes advanced EEG hardware and
signal processing algorithms, employs three self-adhesive, disposable
sensors placed outside the hairline for reliable EEG signal acquisition.
The participants of the SCREENS study were instructed to attach the
device when they went to bed. The ZM uses two proprietary algorithms:
Z-ALG and Z-PLUS. The Z-ALG is utilized for accurate sleep detection,
showcasing its suitability for in-home monitoring[@kaplan2014], while
the Z-PLUS effectively differentiates sleep stages, as evidenced by its
alignment with expert evaluations using PSG data[@wang2015]. In the
current study, we treated all sleep stages as a single category
effectively deducing the output of the ZM to "awake" and "asleep" as the
ability to distinguish sleep stages are not a necessity to derive the
sleep quality parameters of interest and to simplify the learning
process of the models.

@fig-flow illustrates the selection criteria applied to the children's
recordings from the SCREENS study. We included only ZM recordings that
were accompanied by complete accelerometer data and lasted between 7 and
14 hours. Any night when the ZM reported sensor issues was excluded
yielding 585 nights included in the study. The children whose recordings
were considered had an average age of 9.4 years, with a standard
deviation of 2.1. In their raw form, the ZM predictions encompassed
696,779 epochs, each 30 seconds long. Notably, approximately 84% of the
total ZM recording duration was classified as sleep, resulting in an
imbalance of the ZM dataset.

![Flowchart of eligible ZM recording nights included in the
study](visuals/flowchart_of_elligible_nights.pdf){#fig-flow}

Finally, we affirm that the SCREENS study received approval from the
Regional Scientific Committee of Southern Denmark, and all data handling
processes complied with the General Data Protection Regulation (GDPR),
ensuring the ethical and secure management of participant information.

## Data Preprocessing and Feature Extraction

In this study, data processing of the raw accelerometer data began with
a low-pass filtration step using a 4th order Butterworth filter with a 5
Hz cut-off frequency to eliminate high-frequency noise. Following
filtration, data were partitioned into overlapping 2-second intervals,
each successive interval sharing a 50% overlap with the previous one
similar to methods described by Skotte et al.[@skotte_detection_2014].
Any non-wear data was removed using previously described
methods[@skovgaard2023] and data was resampled to 30-second epochs so
every sample classified by the algorithms corresponds to a 30-second
epoch scored during the ZM recordings. Subsequently, we performed a
feature extraction process that yielded a set of 88 features, providing
a robust characterization of the data. Extracted from accelerometer and
temperature signals, these features include temporal elements that use
both lag and lead values, capturing dynamic data trends by incorporating
measurements from preceding and upcoming epochs. Furthermore, inspired
by Walch et al.[@walch2019], we incorporated sensor-independent features
to encapsulate circadian rhythms. These features offer unique insights
not directly discernible from sensor outputs and are meant to
approximate the changing drive of the circadian clock to sleep over the
course of the night (see @fig-sensor-independent). Furthermore, the
feature set was enriched by including signal characteristics, which
encompass vector magnitude, mean crossing rate, skewness, and kurtosis
for each of the x, y, and z dimensions. All features are summarized in
table ??? **in the supplementary matrials**. Subsequently, we merged the
ZM and corresponding accelerometer recordings. Any overlapping time
between the ZM and accelerometer data was treated as 'in-bed' time, with
the remaining time considered 'out-of-bed'. This process yielded a
dataset providing a around the clock temporal view of each participant's
activity and sleep patterns.

![Sensor-independent features of circadian rhythms across two
consecutive nights. A) cosinus feature, B) linear
feature.](visuals/sensor_independent.pdf){#fig-sensor-independent}

In addition to the engineered features, we chose to incorporate the
median-filtered raw predictions from the ZM device into our modeling
process. This decision stemmed from the understanding that children
typically undergo around five to eight sleep cycles per night, with
awakenings most likely occurring at the end of each
cycle[@galland_normal_2012]. Upon examining the raw ZM predictions, we
noted a significant overestimation in the number of awakenings per night
for the children in our study, exceeding what would be expected based on
typical sleep cycle patterns (see @fig-zm-median). In particular, many
of these brief awakenings could be considered as noise, which when
present in the data, can potentially hinder the learning process of
machine learning models by obscuring the underlying patterns that the
models are trying to learn, leading to less accurate predictions.
Consequently, we elected to train and evaluate our models using not only
the raw ZM output, but also versions that were subjected to 5-minute and
10-minute median filters. This approach, by mitigating this noise,
resulted in an anticipated, more age-appropriate count of awakenings per
night, providing a more accurate depiction of children's sleep patterns
(see @tbl-zm_overview).

![The difference in number of awakenings between the raw ZM predictions
vs. 5-minute, and 10-minute median filtered predictions for a random
night. Grey line is the raw predictions, black line is the median
filtered predictions. A: 5-minute median filter on raw ZM predictions,
B: 10-minute median filter on raw ZM
predictions.](visuals/zm_raw_vs_filtered.pdf){#fig-zm-median}

## Algorithms

In this study, we employed two distinct learning strategies to analyze
sleep patterns from thigh-mounted accelerometer data. The first strategy
employed a sequential methodology, using an ensemble of four pairs of
models, with each pair featuring the same algorithm. This sequential
strategy was designed to simplify the prediction task by decomposing it
into two stages: first predicting 'in-bed' time, then 'sleep' time. The
output from the first set of models, which predicted in-bed time, was
subjected to a 5-minute median filter to eliminate leading and trailing
in-bed time blips. This process enabled us to extract a single
continuous time interval that we identified as the sleep period time
window (SPT). The SPT, in turn, served as the input to the second models
in the sequence, enhancing their predictive accuracy for sleep time. We
applied this sequential strategy across all four algorithms, as detailed
below.

1.  Logistic Regression (LREG): Logistic regression served as a simple
    and fast baseline model. However, due to its linear nature, it may
    struggle with capturing complex relationships and non-linear
    patterns present in the accelerometer data.

2.  Decision Tree (TREE): Decision trees are capable of handling
    non-linear patterns and are easily interpretable. However, they are
    prone to overfitting, particularly when dealing with complex
    patterns that require simultaneous consideration of multiple
    features.

3.  Single-layer Feed-forward Neural Network (SNN): Single-layer
    feed-forward neural networks can effectively capture non-linear
    relationships, even with their relatively simple structure. However,
    they tend to be more challenging to interpret compared to simpler
    models. Additionally, careful tuning of the network's architecture
    and training process is required to mitigate the risk of
    overfitting.

4.  XGBoost (XGB): XGBoost is a powerful algorithm known for its ability
    to provide highly accurate predictions and handle complex,
    non-linear patterns in the data. It also incorporates built-in
    methods to prevent overfitting. However, training XGBoost models can
    be computationally intensive, and interpreting the predictions it
    generates can pose challenges.

In parallel, we also employed a multiclass approach using a
bidirectional Long Short-Term Memory (biLSTM)[@hochreiter1997] neural
network. This network, which was designed to predict three distinct
classes: 'out-of-bed-awake', 'in-bed-awake', and 'in-bed-asleep', was
configured with four layers and 128 hidden units per layer. This balance
between model complexity and training efficiency was intended to
facilitate learning of intricate patterns while ensuring feasible
training times. The bidirectional nature of the LSTM enhanced data
interpretation and reduced overfitting by doubling the hidden units at
each time step. The LSTM model used sequences of tensors as input, with
each sequence spanning 10 minutes and a step size of one. As
demonstrated by previous studies such as those by Sano et al.
[@sano2019] and Chen et al. [@chen2021], LSTM models have shown great
promise in sleep detection using accelerometer data, thanks to their
ability to capture complex temporal patterns.

## Model Training

For the models in sequence, we trained four pairs of classification
models. Each pair was designed to distinguish between in-bed/out-of-bed
and asleep/awake states, respectively. The dataset was randomly split
into a training set and a testing set, each containing approximately 50%
of the subjects. The splitting of the data was ensured to not have
samples from the same night simultaneously present in both sets. To
optimize hyperparameters, we performed a 10-fold Monte Carlo
cross-validation on a regular grid, comprising 20 different combinations
of hyperparameters. The F1 score served as the optimization metric. The
best-performing set of hyperparameters was then used to fit the models
to the full training dataset. This approach allowed us to maximize
performance by leveraging all available data. Moreover, after extracting
the in-bed time from the initial sequential models, the imbalance on the
resulting dataset could cause biases during model training, as models
may favor predicting the majority class. To rectify this, we employed
the Synthetic Minority Over-sampling Technique (SMOTE)[@chawla2002].
SMOTE generates new samples by interpolating random samples with their
nearest neighbors. We utilized the themis R package[@themis] to
implement SMOTE, resulting in a balanced distribution of training
samples across both classes.

The biLSTM model was trained to differentiate between three states:
out-of-bed-awake, in-bed-awake, and in-bed-asleep. The data used for
training the biLSTM was randomly divided into training, validation, and
test sets, based on a 50/25/25 split. We ensured that data from the same
night was not present across different sets. The model was trained using
the Adam optimizer, selected for its computational efficiency and
adaptability of the learning rate during training. Given the multiclass
classification task with mutually exclusive classes, we employed the
cross-entropy loss function. To obtain a probability distribution over
the classes, the softmax activation function was applied at the output
layer. We evaluated the model's performance using the F1 score on both
the training and validation sets. We implemented early stopping with a
patience of 3 epochs, halting the training process if there was no
improvement in the validation loss over three consecutive epochs.

## Model Validation

In our study, we utilized standard evaluation metrics to assess the
performance of each model on an epoch-to-epoch basis. These include
accuracy ($accuracy = \frac{TP+TN}{TP+TN+FP+FN}$), sensitivity
($sensitivity = \frac{TP}{TP+FN}$), specificity
($specificity = \frac{TN}{TN+FP}$), precision
($precision = \frac{TP}{TP+FP}$), negative predictive value (NPV,
$NPV = \frac{TN}{TN + FN}$), and F1 score
($F_1 = 2 * \frac{precision * sensitivity}{precision + sensitivity}$).

In the context of our sequential learning strategy, the initial models
were tasked with the binary classification of in-bed vs. out-of-bed. For
this task, we assessed performance using the F1-score, accuracy,
sensitivity, specificity, and precision metrics. The second models in
our sequential learning strategy focused on the binary classification of
asleep vs. awake. For these models, we considered the same metrics, in
addition to the negative predictive rate. The class imbalance in this
case led us to compute the F1 score as an unweighted macro-average.
Additionally, we evaluated the multiclass classifier, biLSTM, using the
same metrics. To do this, we considered the multiclass output as to
binary classifications, where the first was out-of-bed vs the rest and
the second binary classification as in-bed-awake vs in-bed-asleep. To
further illustrate model performance, we provide confusion matrices for
the full dataset, encompassing both in-bed and out-of-bed data. These
matrices report relative counts, column percentages (the proportion of
the true class accurately predicted), and row percentages (the
proportion of predictions correctly classified). We considered both the
in-bed/out-of-bed and awake/asleep scoring tasks as binary
classification problems, designating in-bed and asleep as the positive
labels and out-of-bed and awake as the negative labels in accordance
with previous research[@hjorth2012; @kushida2001].

To assess the performance of our models in deriving sleep quality
parameters, we utilized Bland-Altman plots and Pearson correlations. The
Bland-Altman method was employed specifically to determine the level of
agreement between two measurement techniques. Considering our dataset
contained multiple observations per subject, we integrated a bootstrap
procedure to address this extra source of variability. We calculated the
mean difference (bias) and defined the LOA as the mean difference plus
or minus 1.96 times the standard deviation of these differences. To
ensure our measurements were robust and accounted for intra-subject
variability, we estimated the 95% confidence intervals for both the bias
and the LOA using a bias-corrected and accelerated bootstrap method,
utilizing 10,000 bootstrap replicates. The sleep quality parameteres
included are defined as follews in accordance with the ZM definitions:

1.  Sleep Period Time (SPT) - This refers to the total duration of time
    in bed with the intention to sleep, which is defined as the time
    from the start to the end of the ZM recording.

2.  Total Sleep Time (TST) - This is the time spent asleep within the
    SPT.

3.  Sleep Efficiency (SE) - This is the ratio between TST and SPT,
    representing the proportion of the sleep period that was actually
    spent asleep.

4.  Latency Until Persistent Sleep (LPS) - This metric represents the
    time it takes to transition from wakefulness to sustained sleep. It
    is calculated as the time from the beginning of the ZM recording
    until the first period when 10 out of 12 minutes are scored as
    sleep.

5.  Wake After Sleep Onset (WASO) - This refers to the time spent awake
    after initially falling asleep and before the final awakening. In
    our analysis, a period is counted as 'awake' only if it consists of
    3 or more contiguous 30-second epochs which is also how the ZM
    summarizes WASO.

R version 4.3.0 (2023-04-21)[@R-lang] and the Tidymodels[@tidymodels]
and Tidyverse[@tidyverse] suite of packages were used as the core tools
for model development and analyses. Python version
3.10.6[@10.5555/1593511] and PyTorch[@NEURIPS2019_9015] were used to
implement the biLSTM model. All code used to perform the analysis and
generate the figures in this paper are available in [this
repository](https://github.com/esbenlykke/sleep_study).

# Results

As reported in @tbl-zm_overview the sleep quality parameters derived
from ZM predictions were modified by the implementation of 5-minute and
10-minute median filters. SPT were consistent across raw and filtered
datasets (mean: 9.2 ± 2.1 hours), corresponding to the length of the ZM
recording. TST and SE increased in the filtered data, implying the
filters categorize some wakefulness as sleep. Specifically, TST
increased from a raw mean of 7.7 ± 1.9 hours to 8.1 ± 2.0 hours
(5-minute filter) and 8.2 ± 2.1 hours (10-minute filter), while SE rose
from 82.6 ± 12.0% to 86.4 ± 12.7% and 87.5 ± 12.9% respectively. LPS
also elevated, suggesting the filter smooths out brief awakenings at
sleep onset, leading to a prolonged time to persistent sleep. A
significant change was seen in WASO, which dropped from 39.0 ± 33.6
minutes in raw data to 30.6 ± 46.8 minutes and 22.3 ± 55.4 minutes in
the 5-minute and 10-minute filtered data, respectively. The number of
awakenings was also considerably reduced with the application of
filters. In the raw data, the average number of awakenings was 34.46 ±
11.33 per night, which reduced to 4.43 ± 3.26 and 1.95 ± 2.01 for the
5-minute and 10-minute filtered data sets respectively.

```{r}
#| echo: false
#| message: false
#| label: tbl-zm_overview
#| tbl-cap: Overview of characteristics of the ZM sleep quality summaries per night. Values are represented as mean (SD).

source("code/table_overview_ZM_stats.R")

tbl_overview
```

## Performance on Epoch-to-Epoch Basis

The epoch-to-epoch evaluation of predicting in-bed time is outlined in
@tbl-in_bed_performance, and demonstrates practically equivalent
performance across all model types. The F1 score ranges from 94.4%
(Decision Tree) to 95.4% (XGBoost), while accuracy ranges from 95.3%
(Decision Tree) to 96.1% (XGBoost). Sensitivity, Precision, and
Specificity also demonstrate consistent results across the models. The
XGBoost model, despite recording the highest metrics with an F1 score of
95.4% and accuracy of 96.1%, outpaced the others only marginally.

```{r}
#| echo: false
#| message: false
#| label: tbl-in_bed_performance
#| tbl-cap: In-bed performance metrics

source("code/table_in_bed_performance.R")

tbl_in_bed
```

@tbl-sleep_performance details the performance of all sequential model
types on raw and median-filtered (5 and 10 minute) ZM predictions for
sleep/wake classification. For raw ZM predictions, the F1 scores, which
are unweighted macro averages, range from 65.6% (biLSTM) to 76.2%
(XGBoost). All models perform comparably, but the low specificity values
(62.5% to 70.9%) suggest difficulty in correctly classifying awake
epochs. Applying a 5-minute median filter improves the performance
metrics. The XGBoost model tops the charts with an F1 score of 79.2% and
NPV of 74.0%. However, specificity still remains low, with values
between 54.7% (XGBoost) and 74.8% (Logistic Regression) across all
models. With a 10-minute median filter, the metrics improve further. The
XGBoost model still leads with an F1 score of 80.9% and an NPV of 75.9%.
But, specificity remains low, ranging from 57.5% (Decision Tree) to
76.4% (Logistic Regression) across all models.

```{r}
#| echo: false
#| message: false
#| tbl-colwidths: auto
#| label: tbl-sleep_performance
#| tbl-cap: Performance of the sleep/wake classification of the sequential models.

source("code/table_sleep_performance.R")

tbl_sleep_performance
```

A complete set of confusion matrices generated from data both containing
the out-of-bed and in-bed time are presented in @fig-conf_mat. The
figure shows favorable epoch-to-epoch performance across all sequential
models, however, it is evident that the biLSTM is less successful in
classifying the in-bed-awake class which cannot be deduced from the
confusion matrices from the sequential models.

![Confusion matrices for binary sleep prediction. The middle of each
tile is the normalized count (overall percentage) and, beneath it, the
count. The bottom number is the column percentage (target). At the right
side of each tile is the row percentage (prediction). i) decision tree,
ii) logistic regression, iii) feed-forward neural net, iv) XGBoost, and
v) biLSTM.](visuals/all_conf_mats.pdf){#fig-conf_mat}

## Evaluation of Sleep Quality Parameters

@tbl-ba_cor presents a comparative analysis of the included models used
to predict various sleep quality parameters (SPT, TST, SE, LPS, WASO)
using the 5-minute median filtered ZM predictions. To see the full table
including models developed from raw ZM predictions and 10-minute median
filtered ZM predictions, see **\[SUPP. MAT.\]**. In terms of bias, the
decision tree model consistently underestimated SPT, TST, and SE, and
overestimated LPS and WASO in comparison to ZM. The logistic regression
model had similar trends, with more pronounced underestimation in TST
and overestimation in LPS. The feed-forward neural network also
exhibited similar bias as the decision tree and the logistic regression
models, but with a higher overestimation in WASO. On the other hand, the
XGBoost model showed least bias among all, especially in its 5-minute
median predictions. Considering LOA, the decision tree had higher
variability across different sleep quality parameters and filtering
techniques, particularly for LPS and WASO, which indicates lower
agreement with ZM. Other models had comparable LOA but with notable
exceptions. For example, TST LOA for the logistic regression model was
particularly wide in the 5-minute median predictions. Correlation-wise,
the pearson coefficient, revealed that the XGBoost model consistently
had the highest correlation with ZM across all sleep quality parameters
and filtering methods Notably, the XGBoost's 5-minute median predictions
showed the strongest correlation (0.66) for TST among all models and
filtering techniques.

```{r}
#| echo: false
#| message: false
#| label: tbl-ba_cor
#| tbl-cap: Summary of Bias, Limits of Agreement, and Pearson Correlation for various Sleep Parameter Predictions (SPT, TST, SE, LPS, WASO) using different Machine Learning Models (Decision Tree, Logistic Regression, Feed-Forward Neural Net, XGBoost) with Raw ZM Predictions, 5-Min and 10-Min Median as predictors. Each value is provided with its 95% Confidence Interval (CI).

source("code/table_ba_cor_statistics.R")

short_tbl_ba_cor

```

@fig-xgb_ba_cor shows the agreement between the XGBoost model, trained
on 5-minute median filtered ZM predictions, and the 5-minute
median-smoothed ZM-derived sleep quality parameters. The Bland-Altman
plot for the SPT and TST reveals a significant level of agreement with
the ZM, as evidenced by a bias close to zero. Interestingly, a portion
of the data points are located near the zero line indicating perfect
agreement. However, the presence of extreme outliers widens the limits
of agreement (LOA). The scatterplot for SPT also demonstrates a positive
trend, indicating a moderate linear correlation between the XGBoost
model and the ZM-derived sleep quality parameters. The bias and LOA for
TST are comparable to those observed for SPT, indicating a consistent
level of agreement between the two methods. The scatterplot for TST also
shows a slightly higher correlation, primarily driven by the absence of
extreme outliers.Furthermore, the remaining three sleep quality
parameters, SE, LPS, and WASO, exhibit heteroscedasticity in contrast to
SPT and TST. This outcome is expected as achieving 100% sleep efficiency
is relatively rare, resulting in less disagreement between the methods
as values approach the upper limit. However, as sleep efficiency
decreases, the potential for discrepancies and differing interpretations
between the methods increases, leading to greater heteroscedasticity. A
moderate positive linear correlation is observed between the XGBoost
model and ZM-derived sleep quality parameters for SE, however, a poor
correlation is observed for LPS and WASO, suggesting less agreement
between the two methods for these parameters. Similar plots for all
models are available in the **supplementary materials (75 plots. Jeg ved
ikke om nogen orker at bladre igennem dem).**

![Comparison of sleep quality parameters derived from the XGBoost model
trained on the 5-minute smoothed ZM predictions. The left column
displays Bland-Altman plots. Dashed lines represent the bias (the
average difference between the two measurements) and LOA, with the 95%
confidence intervals represented as the grayed areas. The right column
displays scatter plots of XGBoost-derived vs ZM-derived sleep quality
parameters. The dashed line represents the identity line, while the
full-drawn line represents the best linear fit. Pearson's correlations
are annotated in the upper left
corner.](visuals/median_5_xgboost_ba_cor.pdf){#fig-xgb_ba_cor}

# Discussion

In an effort to mature the methods for estimating sleep from thigh-worn
accelerometers, we evaluated various models for predicting in-bed and
sleep time and their derived sleep quality parameters. Furthermore, we
trained and evaluated the models using raw and median-filtered gold
standard predictions from the ZM. In general, all sequential models
performed well at predicting in-bed time. More challenging was it to
distinguish wake from sleep on the extracted in-bed time, however, the
performance of the sequential models were enhanced by the application of
median filters. Moreover, even though the multiclass biLSTM showed good
performance across F1 score, precision and NPV, the derived sleep
quality parameters were not on par with the XGBoost model which
demonstrated the highest performance metrics across all evaluations,
including epoch-to-epoch prediction and sleep quality parameter
derivatives. Despite this, all sequential models showed low specificity
values, indicating difficulty in correctly classifying awake epochs
during time in bed. The application of 5-minute and 10-minute median
filters improved the performance metrics of all models, with the XGBoost
model consistently leading. The median filters increased total sleep
time and sleep efficiency, while reducing wake after sleep onset and the
number of awakenings. The XGBoost model also showed the least bias and
highest correlation with ZM across sleep quality parameters.

Limited research exists regarding the epoch-to-epoch effectiveness of
classifying in-bed time based on data from thigh-worn accelerometers.
Nevertheless, Carlson and colleagues provided compelling insights. They
demonstrated that a third-party algorithm, "ProcessingPal," and a
proprietary one, "CREA," achieved accuracies of 91% and 86%
respectively. These algorithms, evaluated against self-reported
measures[@carlson2021], produced F1 scores as high as 95% and 96%. These
figures are consistent with the performance of our sequential models,
which also achieved F1 scores and accuracy scores exceeding 95% in
identifying in-bed time. In our study, in-bed time is equated with SPT.
All models, with the exception of XGBoost, underestimated SPT. The
biLSTM model showed the greatest underestimation, with a bias of -36
minutes, reflecting trends observed in previous research. Winkler et al.
developed an algorithm that, despite a strong correlation (Pearson
correlation coefficient = .67) between their algorithmic results and
diary-recorded waking times, overestimated waking wear time by more than
30 minutes, resulting in an underestimation of in-bed
time[@winkler2016]. This trend was further confirmed when Inan-Eroglu et
al. examined Winkler et al.'s algorithm, revealing a underestimation of
9.8 minutes in bed time compared to self-reported
measures[@inan-eroglu2021]. In contrast, a study by van der Berg et al.
reported a slight underestimation of in-bed time. They employed a unique
approach with their algorithm, which relied on quantifying the number
and duration of sedentary periods to determine time in bed, and active
periods (standing or stepping) to identify wake times[@vanderberg2016].
Finally. it is important to note that high predictive performance in
determining in-bed time does not necessarily translate to accurate
predictions of broader sleep quality parameters. The crucial task of
detecting awake periods during in-bed time, a key factor in assessing
sleep quality, may not be effectively captured by in-bed time
predictions alone. Indeed, underestimating in-bed time could result in
overestimating waking time during in-bed time. Furthermore, the
distinction between actual sleep and time spent in bed, often overlooked
but vital in sleep research, is critical for a comprehensive
understanding of sleep quality.

To the best of our knowledge, Johansson and
colleagues[@johansson_development_2023] are the only researchers who
have reported epoch-to-epoch performance metrics for sleep scoring using
thigh-worn accelerometers, beyond just "waking time" and "in-bed time."
They achieved a mean sensitivity of 0.84, specificity of 0.55, and
accuracy of 0.80, using a single-night evaluation dataset of 71
subjects. Despite our models achieving a sensitivity above 97%, they,
like Johansson et al.'s algorithm, struggled with detecting in-bed awake
epochs. This is reflected in the low specificity scores, ranging from
54.7% to 76.4%, reported in our study. The challenge of low specificity
is not unique to methods using data collected from thigh-worn devices.
Conley et al.'s meta-analysis[@conley2019] reported similar findings
when estimating sleep using wrist-worn accelerometers among healthy
adults, with a mean sensitivity, accuracy, and specificity of 0.89,
0.88, and 0.53, respectively. Furthermore, Patterson and
colleagues[@patterson_40_2023] recently summarized the performance of
various heuristic algorithms, machine learning, and deep learning models
used to predict sleep. They found the mean sensitivity and specificity
to be 93% (SD = 2.8) and 60% (SD = 11.1) respectively. These findings
underscore the challenge of automating the detection of in-bed awake
periods. Interestingly, despite low specificity values for most of our
models and configurations, we observed an overestimation of LPS and
WASO, contrasting with most previous research[@conley2019;
@palotti2019]. This overestimation of wake epochs is evident from the
low NPV scores, indicating that only a small proportion of the wake
predictions are actually correct. This discrepancy may be driven by the
SMOTE process used to balance the dataset. If the synthetic "wake"
samples created by SMOTE are not representative of the true "wake" data,
the models might learn to incorrectly classify certain "sleep" epochs as
"wake". This could lead to an overestimation of LPS and WASO, as the
models are incorrectly identifying more periods of wakefulness during
the sleep period.

The use of the SMOTE technique likely improved the performance of our
models by addressing the class imbalance in our data. However, this
technique also introduced synthetic"wake" samples that may not be fully
representative of true wake data. This could potentially lead some
models to overestimate the wake class. Interestingly, the biLSTM model,
which was not trained on SMOTE-processed data, was the only one to
overestimate TST and SE. On the other hand, the XGBoost model, which was
trained on data subjected to the SMOTE process, was able to handle the
synthetic "wake" samples better than the other models, and it did not
overestimate TST to the same degree. The Bland-Altman statistics for the
XGBoost model trained on the 5-minute median filtered ZM predictions
showed a mean difference of -7 minutes for TST and -1.1% for SE, with
limits of agreement ranging from -95.5 to 81.4 minutes and from -15.6%
to 13.3% respectively. This suggests that the XGBoost model was able to
maintain a balance between sensitivity and specificity, and it was not
overly influenced by the synthetic "wake" samples. The XGBoost model's
success with the SMOTE dataset may be due to its ability to handle
non-representative synthetic samples. XGBoost's gradient boosting
mechanism allows it to iteratively learn from the errors of previous
models, which can help it to better distinguish between true wake data
and synthetic wake samples created by SMOTE. This iterative learning
process could make XGBoost more robust to the inaccuracies introduced by
the synthetic samples, leading to better overall performance.

Typically, sleep detection methods are applied in two contexts: either
to night recordings or to 24-hour recordings. In night recordings, it is
possible to derive sleep quality parameters like SE and LPS because the
SPT is already known because it is inferred from the length of the
recording[@conley2019; @patterson_40_2023]. On the other hand, when
sleep detection methods are applied to 24-hour recordings, most methods
do not have the ability to infer the SPT with sleep
diaries[@girschik2012]. Consequently, these methods are unable to
generate certain sleep quality parameters that rely on the
SPT[@doherty2017; @anderson2014]. To overcome this limitation, we have
incorporated models that can differentiate between in-bed awake time and
in-bed asleep from out-bed awake time over a 24-hour recording. This
approach allows our models to estimate all commonly used sleep quality
parameters. Van Hees et al. [@van_hees_estimating_2018] have proposed an
algorithm to determine SPT from data collected by wrist-worn devices.
This algorithm was recently validated by Plekhanova and her team
[@plekhanova2023]. By combining this algorithm with other methods,
further sleep quality parameters can be inferred based on the identified
SPT. Van Hees et al.[@van_hees_estimating_2018] reported good agreements
and low mean differences compared to self-report and PSG on SPT,
findings later confirmed by Plekhanova and colleagues. However, they
also observed poor agreement with LPS and Wake After Sleep Onset (WASO).
They found low reliability with PSG, indicating difficulties in
detecting wakefulness during in-bed time. These challenges parallel
those we experienced in our study.

In our evaluation of sleep quality parameters, we found that LPS had the
largest mean error relative to absolute time allocated to LPS. This
suggests that the initial epochs of Sleep Period Time (SPT) are
particularly challenging to classify correctly. This is also supported
by the poor Pearson correlations between LPS derived from model
predictions and the ZM. The XGBoost model, which was the best performer
among all models, overestimated LPS by an average of 26.4 minutes for
models trained on raw ZM predictions, 28.5 minutes for models trained on
5-minute filtered ZM predictions, and 34.5 minutes for models trained on
10-minute filtered ZM predictions. This level of discrepancy is
comparable to the mean error of sleep latency of 23 minutes reported by
Johansson et al.[@johansson_development_2023]. Johansson et al. suggest
that the discrepancy with the gold standard is likely due to the
multifaceted nature of the sleep state, which is a complex physiological
process. Short awakenings or sleep episodes may not necessarily
correspond to noticeable changes in thigh movement, making them
difficult to detect and accurately classify. These results align with
several methods for wrist-worn devices reviewed by Conley and
colleagues[@conley2019]. They reported correlations between
accelerometer and PSG sleep onset latency (equivalent to LPS) from 10
studies with a mean correlation of 0.2 (ranging from -0.69 to 0.69),
indicating the inherent difficulty in estimating this parameter using
accelerometry alone.

Our study's XGBoost model demonstrated relatively narrower limits of
agreements (LOAs) for TST, SE, and WASO, with ranges of -95.5 to 81.4
min, -15.6 to 13.3%, and -83.4 to 81.7 min, respectively when compared
with other models such as the Van Hees algorithm[@hees2015], Oakley rsc
(rescored)[@palotti2019], and LSTM-50[@palotti2019] evaluated in the
Patterson et al. study[@patterson_40_2023]. Furthermore, comparing the
LOAs between our XGBoost model and the algorithm developed for
thigh-worn devices by Johansson et al.
study[@johansson_development_2023], our XGBoost model showed narrower
LOAs for TST , SE, LPS , and WASO, but not SPT. Generally, all methods,
both from this study and from the reviewed literature, exibit wide LOAs
sugesting that there is high variability in the derevide sleep quality
parameters. These findings imply that the current methods, are only
reasonbly reliable for assessing sleep quality parameters at a group
level. However, caution should be exercised when applying the models and
methods to individual-level sleep assessments. Therefore, further
improvements and refinements are needed to enhance the precision and
reliability of these models for individual sleep assessments.

In this study, we used the ZM as the reference method, rather than PSG,
which is considered the gold standard for sleep measurement. This choice
may contribute to discrepancies between our models and the ZM, as
without a true gold standard, it's difficult to determine the source of
disagreement. However, we believe that the use of ZM, which allows for
multiple consecutive nights of recording, is valuable. This approach
captures intra-individual variances in sleep, which is impractical with
PSG. It also enabled us to include more nights in our study typically
compared to those relying on PSG. For instance, the widely used
Newcastle dataset[@hees2015] only contains data from 28 participants.
However, upon examining the ZM outputs, we found that the raw
predictions were not optimal for developing machine learning models due
to a seemingly low signal-to-noise ratio (see @fig-zm-median). The ZM
itself mitigates this issue by applying certain filtering processes when
generating sleep quality parameters. For example, epochs contributing to
WASO must be in contiguous epochs of 3, and sleep only counts towards
sleep quality parameters if 10 out of 12 minutes are scored as sleep. To
improve the prospect of our machine learning algorithms, we applied
median filters to the ZM raw predictions. This did in fact alter the
derived sleep quality parameters. Notably, the mean WASO decreased from
39 minutes in the raw predictions to 30.6 minutes in the 5-minute median
filtered predictions, and further decreased to 22.3 minutes in the
10-minute filtered predictions. The application of 5-minute and
10-minute median filters also led to increases in TST, SE, and LPS. This
suggests that the filters may categorize some instances of wakefulness
as sleep and smooth out brief awakenings. Despite these changes, the
overall sleep quality profile derived from the median-filtered
predictions is still comparable to that from the raw predictions,
justifying our approach.

The study boasts several strengths, including the capacity to
distinguish in-bed awake and asleep from out-of-bed, thereby allowing
for the extraction of vital sleep quality parameters. Furthermore, the
research benefits from evaluating multiple nights per subject, providing
valuable information into intra-subject sleep variability. However,
certain limitations exist. The use of ZM, which isn't recognized as a
gold standard, could potentially compromise our findings' validity.
Future research could consider using PSG as a reference for methods
similar to ours, despite its limitations, for a more accurate
comparison. Moreover, our models weren't validated using an external
dataset, a process that would have showcased their broader
applicability. Hence, our conclusions remain confined primarily to
children.

In conclusion, our study contributes to the ongoing efforts to improve
sleep estimation methods using thigh-worn accelerometers. We evaluated
different machine learning models for predicting in-bed and sleep times
and their corresponding sleep quality parameters. While the sequential
models generally demonstrated excellent performance in predicting in-bed
time, they faced challenges in accurately distinguishing between sleep
and wake epochs during in-bed time. Among all models and configurations
evaluated, the XGBoost model exhibited superior performance across all
performance metrics, including epoch-to-epoch predictions and sleep
quality parameter derivatives. Our research also highlighted the current
limitations of sleep detection methods, such as challenges in
effectively detecting wake periods during in-bed time and the need for
further improvements to increase the precision of individual sleep
assessments. We believe our work lays the groundwork for future research
to further refine and improve the performance of these models,
contributing to a more precise and accurate evaluation of sleep patterns
and quality using thigh-worn accelerometers.

\newpage

# References