### 2023-08-01 manuscript/supp_mat/supp.qmd
manus

---
title: "Supplmentary Material for DOI:"
date: last-modified
format:
  pdf
---

Table of features? Er det nødvendigt?

```{r}
#| echo: false
#| message: false
#| label: tbl-ba_cor
#| tbl-cap: Summary of Bias, Limits of Agreement (LOA), and Pearson Correlation for various Sleep Parameter Predictions (SPT, TST, SE, LPS, WASO) using different Machine Learning Models (Decision Tree, Logistic Regression, Feed-Forward Neural Net, XGBoost) with Raw ZM Predictions, 5-Min and 10-Min Median as predictors. Each value is provided with its 95% Confidence Interval (CI).

source(here::here("manuscript/code/table_ba_cor_statistics.R"))

tbl_ba_cor
```

The analysis of precision-recall and ROC curves across different models
and ZM prediction types shows varying performance. In terms of
precision-recall AUC, the Decision Tree model consistently outperforms
others, indicating its superior predictive accuracy (see
@fig-pr_curves). Conversely, the Neural Network model generally shows
weaker performance. However, for ROC AUC, the XGBoost model consistently
excels across all data types, indicating a strong ability to
differentiate between classes, while the Neural Network model tends to
underperform (see @fig-roc_curves). The F-measure (F1 score) shows
variable performance across different configurations but generally, the
Decision Tree model yields higher scores.

![Precision-recall curves of the models evaluated across the different ZM predictions, including raw ZM predictions, as well as 5-minute and 10-minute median smoothing of the ZM raw predictions. The x-axis of the plot represents the proportion of true wake epochs that were correctly classified as wake, while the y-axis represents the proportion of all epochs labeled as wake by the classifier that were classified correctly. The area under the curve values are displayed as color-coded text in the plot indicate the area under the Precision-Recall curve for each model and condition.](../visuals/plot_sleep_pr.pdf){#fig-pr_curves}

![Receiver operating characteristic curves of the models evaluated across the different ZM predictions, including raw ZM predictions, as well as 5-minute and 10-minute median smoothing of the ZM raw predictions. The x-axis of the plot represents the proportion of true asleep epochs that were incorrectly classified as awake, while the y-axis represents the proportion of all epochs labeled as awake by the classifier that were correctly classified. The area under the curve values displayed are displayed as color-coded text in the plot to indicate the area under the receiver operating characteristic curve for each model and condition.](../visuals/plot_sleep_roc.pdf){#fig-roc_curves}

Plots similar to figure 5 for all model and filtering combinations (75 plots in total) are available upon request.

<!-- ![Decision tree, raw](../visuals/raw_decision_tree_ba_cor.pdf){#tbl-raw_dc} -->
<!-- ![Decision tree, median_5](../visuals/median_5_decision_tree_ba_cor.pdf){#tbl-m5_dc} -->
<!-- ![Decision tree, median_10](../visuals/median_10_decision_tree_ba_cor.pdf){#tbl-m10_dc} -->

<!-- ![Logistic regression, raw](../visuals/raw_logistic_regression_ba_cor.pdf) -->
<!-- ![Logistic regression, median_5](../visuals/median_5_logistic_regression_ba_cor.pdf) -->
<!-- ![Logistic regression, median_10](../visuals/median_10_logistic_regression_ba_cor.pdf) -->

<!-- ![Neural_network, raw](../visuals/raw_neural_network_ba_cor.pdf) -->
<!-- ![Neural_network, median_5](../visuals/median_5_neural_network_ba_cor.pdf) -->
<!-- ![Neural_network, median_10](../visuals/median_10_neural_network_ba_cor.pdf) -->

<!-- ![XGBoost, raw](../visuals/raw_xgboost_ba_cor.pdf) -->
<!-- ![XGBoost, median_5](../visuals/median_5_xgboost_ba_cor.pdf) -->
<!-- ![XGBoost, median_10](../visuals/median_10_xgboost_ba_cor.pdf) -->

<!-- ![biLSTM, raw](../visuals/raw_biLSTM_ba_cor.pdf) -->
<!-- ![biLSTM, median_5](../visuals/median_5_biLSTM_ba_cor.pdf) -->
<!-- ![biLSTM, median_10](../visuals/median_10_biLSTM_ba_cor.pdf) -->